### 2016-07-28 experiments/e2/writeup/print_info_on_design.Rmd

```{r}
# print info on design, number of cells
cat(paste('number of subjects = ',30,'\n',
          '---','\n',
          'number of trials per subject = ',256,'\n', 
          '---','\n',          
          'number of unique cells in the design = Item(4) * Number(2) * Vagueness(2) * Order(2) * Quantity(2) = ',4*2*2*2*2,'\n',
          'number of times a subject saw a unique cell in the design = 256 / 64 = ', 256/64,'\n',
          '---','\n',
          'number of higher-level cells in the design (abstracting over Order and Quantity) = Item(4) * Number(2) * Vagueness(2) = ',4*2*2,'\n',
          'number of times a subject saw a higher level cell in the design =  = 256 / 16 = ', 4*2*2, '\n',
          sep=''))
```