### 2016-07-28 experiments/e2/writeup/show_compression_of_ratios.Rmd

```{r, fig.width=10, fig.height=7, echo=F}
#  Show compression of ratios of numbers of dots in the items
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)

plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')

points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
text(y=1.15, x=0.1, labels="mean (ratio min/med, ratio med/max)", col='red')
points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
axis(1, at=seq(0,1,0.1))

```