### 2016-07-29 experiments/e2/archive/plot_boxplot_subjects_and_items.Rmd

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r, fig.width=20, fig.height=5}
# boxplot subjects and items

load("dd.Rda")
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))

# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value")

# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```