### 2016-07-29 experiments/e2/archive/table_raw_mean_RT_tables.Rmd

## Raw mean RT tables

```{r, results='asis', eval=TRUE, cache=FALSE}
var1 = ddply(dat, .(Vagueness), summarise, RT=mean(RT, na.rm=TRUE) )
print(xtable(var1, digits=0),type='html', include.rownames=F)
```

```{r, results='asis', eval=TRUE, cache=FALSE}
var2=ddply(dat, .(Number), summarise, RT=mean(RT,na.rm=TRUE))
print(xtable(var2, digits=0),type='html', include.rownames=F)
```

```{r, results='asis', eval=TRUE, cache=FALSE}
var3=ddply(dat, .(Vagueness,Number), summarise, RT=mean(RT, na.rm=TRUE))
print(xtable(var3, digits=0),type='html', include.rownames=F)
```

```{r, results='asis', eval=TRUE, cache=FALSE}
var4=ddply(dat, .(f_Itm), summarise, RT=mean(RT, na.rm=TRUE))
print(xtable(var3, digits=0),type='html', include.rownames=F)
```