### 2016-07-29 experiments/e2/writeup/e2.Rmd
simplifying the directory by adding function code into the main .Rmd file

---
title: "Experiment 2 (of 4)"
author: "Matt Green"
output: 
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---

```{r loadKnitr}
suppressPackageStartupMessages(library(knitr))
```

```{r loadLibs}
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(latticeExtra))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(beepr))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(LMERConvenienceFunctions)) # for perSubjectTrim
suppressPackageStartupMessages(library(languageR))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(rms))                      # for varclus
suppressPackageStartupMessages(library(languageR))
```

```{r CIfunctions}
## Summarizes data.
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  # This is does the summary; it's not easy to understand...
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun= function(xx, col, na.rm) {
                   c( N    = length2(xx[,col], na.rm=na.rm),
                      mean = mean   (xx[,col], na.rm=na.rm),
                      sd   = sd     (xx[,col], na.rm=na.rm)
                   )
                 },
                 measurevar,
                 na.rm
  )
  # Rename the "mean" column    
  datac <- rename(datac, c("mean"=measurevar))
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  return(datac)
}
## Summarizes data, handling within-subjects variables by removing inter-subject variability.
## It will still work if there are no within-S variables.
## Gives count, un-normed mean, normed mean (with same between-group mean),
##   standard deviation, standard error of the mean, and confidence interval.
## If there are within-subject variables, calculate adjusted values using method from Morey (2008).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   betweenvars: a vector containing names of columns that are between-subjects variables
##   withinvars: a vector containing names of columns that are within-subjects variables
##   idvar: the name of a column that identifies each subject (or matched subjects)
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySEwithin <- function(data=NULL, measurevar, betweenvars=NULL, withinvars=NULL,
                            idvar=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {
  # Ensure that the betweenvars and withinvars are factors
  factorvars <- vapply(data[, c(betweenvars, withinvars), drop=FALSE],
                       FUN=is.factor, FUN.VALUE=logical(1))
  if (!all(factorvars)) {
    nonfactorvars <- names(factorvars)[!factorvars]
    message("Automatically converting the following non-factors to factors: ",
            paste(nonfactorvars, collapse = ", "))
    data[nonfactorvars] <- lapply(data[nonfactorvars], factor)
  }
  # Get the means from the un-normed data
  datac <- summarySE(data, measurevar, groupvars=c(betweenvars, withinvars),
                     na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
  # Drop all the unused columns (these will be calculated with normed data)
  datac$sd <- NULL
  datac$se <- NULL
  datac$ci <- NULL
  # Norm each subject's data
  ndata <- normDataWithin(data, idvar, measurevar, betweenvars, na.rm, .drop=.drop)
  # This is the name of the new column
  measurevar_n <- paste(measurevar, "_norm", sep="")
  # Collapse the normed data - now we can treat between and within vars the same
  ndatac <- summarySE(ndata, measurevar_n, groupvars=c(betweenvars, withinvars),
                      na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
  # Apply correction from Morey (2008) to the standard error and confidence interval
  #  Get the product of the number of conditions of within-S variables
  nWithinGroups    <- prod(vapply(ndatac[,withinvars, drop=FALSE], FUN=nlevels,
                                  FUN.VALUE=numeric(1)))
  correctionFactor <- sqrt( nWithinGroups / (nWithinGroups-1) )
  # Apply the correction factor
  ndatac$sd <- ndatac$sd * correctionFactor
  ndatac$se <- ndatac$se * correctionFactor
  ndatac$ci <- ndatac$ci * correctionFactor
  # Combine the un-normed means with the normed results
  merge(datac, ndatac)
}
## Norms the data within specified groups in a data frame; it normalizes each
## subject (identified by idvar) so that they have the same mean, within each group
## specified by betweenvars.
##   data: a data frame.
##   idvar: the name of a column that identifies each subject (or matched subjects)
##   measurevar: the name of a column that contains the variable to be summariezed
##   betweenvars: a vector containing names of columns that are between-subjects variables
##   na.rm: a boolean that indicates whether to ignore NA's
normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
                           na.rm=FALSE, .drop=TRUE) {
  require(plyr)
  # Measure var on left, idvar + between vars on right of formula.
  data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
                         .fun = function(xx, col, na.rm) {
                           c(subjMean = mean(xx[,col], na.rm=na.rm))
                         },
                         measurevar,
                         na.rm
  )
  # Put the subject means with original data
  data <- merge(data, data.subjMean)
  # Get the normalized data in a new column
  measureNormedVar <- paste(measurevar, "_norm", sep="")
  data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
    mean(data[,measurevar], na.rm=na.rm)
  # Remove this subject mean column
  data$subjMean <- NULL
  return(data)
}
```

```{r print_info_on_design}
cat(paste('number of subjects = ',30,'\n',
          '---','\n',
          'number of trials per subject = ',256,'\n', 
          '---','\n',          
          'number of unique cells in the design = Item(4) * Number(2) * Vagueness(2) * Order(2) * Quantity(2) = ',4*2*2*2*2,'\n',
          'number of times a subject saw a unique cell in the design = 256 / 64 = ', 256/64,'\n',
          '---','\n',
          'number of higher-level cells in the design (abstracting over Order and Quantity) = Item(4) * Number(2) * Vagueness(2) = ',4*2*2,'\n',
          'number of times a subject saw a higher level cell in the design =  = 256 / 16 = ', 4*2*2, '\n',
          sep=''))
```

# Distributional stuff, including transformations.

```{r transformations_table, results='asis'}
x=445
lambda = c(-2,-1,-0.5, 0,0.5,1,2)

xd=data.frame(
  lambda = c(
    lambda[1],
    lambda[2],
    lambda[3],
    lambda[4],
    lambda[5],
    lambda[6],
    lambda[7]
  ),
  x=c(    
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x)
     ),
  conventional=c(
    '$$y=\\frac{1}{x^2}$$', 
    '$$y=\\frac{1}{x}$$', 
    '$$y=\\frac{1}{\\sqrt{x}}$$', 
    '$$y=log_e(x)$$', 
    '$$y=\\sqrt{x}$$', 
    '$$y=x$$', 
    '$$y=x^2$$'
    ),
  power=c(
    '$$y=x^{-2}$$', 
    '$$y=x^{-1}$$', 
    '$$y=x^{-0.5}$$', 
    '$$y=log_e(x)$$', 
    '$$y=x^{0.5}$$', 
    '$$y=x^{1}$$', 
    '$$y=x^{2}$$'
    ),
  y=c(     
    paste("y = ", sep="", format(round(x^lambda[1],6), sc=F) ),
    paste("y = ", sep="", round(x^lambda[2],3) ),
    paste("y = ", sep="", round(x^lambda[3],3) ),
    paste("y = ", sep="", round(log(x),3) ),
    paste("y = ", sep="", round(x^lambda[5],2) ),
    paste("y = ", sep="", round(x^lambda[6],6) ),
    paste("y = ", sep="", round(x^lambda[7],6) )
  ),
  back = c(
    '$$x=y^{\\frac{1}{-2}}$$',
    '$$x=y^{\\frac{1}{-1}}$$', 
    '$$x=y^{\\frac{1}{-0.5}}$$', 
    '$$x=exp(y)$$',  
    '$$x=y^{\\frac{1}{0.5}}$$', 
    '$$x=y^{\\frac{1}{1}}$$',  
    '$$x=y^{\\frac{1}{2}}$$'
  ),
  back_lambda=c(
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=exp(y)$$',  
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$'
  ),
  y_back=c(  
    paste( 'x = ', (x^lambda[1]) ^ {1/lambda[1]} ,sep=""),
    paste( 'x = ', (x^lambda[2]) ^ {1/lambda[2]} ,sep=""),
    paste( 'x = ', (x^lambda[3]) ^ {1/lambda[3]} ,sep=""),
    paste( 'x = ', exp(log(x)), sep=""),
    paste( 'x = ', (x^lambda[5]) ^ {1/lambda[5]} ,sep=""),
    paste( 'x = ', (x^lambda[6]) ^ {1/lambda[6]} ,sep=""),
    paste( 'x = ', (x^lambda[7]) ^ {1/lambda[7]} ,sep="")
  )
)

print(xtable(xd, digits=1, align=c(rep('l', times=ncol(xd)+1))), type='html', include.rownames=F, digits=1)
```

Load the file dat.Rda if it exists, else Get the data if file dat.Rda does not already exist.

```{r, get_the_data}
# if the output file dat.Rda already exists then load it, else do data wrangling
if( file.exists('dat.Rda') && file.exists('dd.Rda') ) {
	load('dat.Rda'); load('dd.Rda')
	} else {
	gatherData = function(number_of_valid_subjects) {
		data_dir <- '../experimentCode/output/'
		column.headers.df <-  head( read.table(
			paste(data_dir,'subject1.data',sep=''),
			header=TRUE),0)
		gathered.data <- column.headers.df
		for (subject in 1:number_of_valid_subjects) {
			current.filename <- paste(data_dir, 'subject', subject, '.data', sep='')
			current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
			gathered.data <- rbind(gathered.data, current.data)
		}
		return(gathered.data)
	} # end of gatherData function
	classifyResponse = function(dat) {
		# what were they expected to respond?
		dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', dat$Quantity, ':Item', dat$Item, sep=''))
		dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con1:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con1:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con1:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con1:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con1:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con1:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con1:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con1:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con1:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con1:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con1:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con1:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con1:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con1:Quan2:Item4', 'Extr_Num']      <- 36
		dat[dat$crossed=='Con2:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con2:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con2:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con2:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con2:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con2:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con2:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con2:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con2:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con2:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con2:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con2:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con2:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con2:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con2:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con2:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con2:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con2:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con2:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con2:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con2:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con2:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con2:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con2:Quan2:Item4', 'Extr_Num']      <- 36
		dat[dat$crossed=='Con3:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con3:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con3:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con3:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con3:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con3:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con3:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con3:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con3:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con3:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con3:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con3:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con3:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con3:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con3:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con3:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con3:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con3:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con3:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con3:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con3:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con3:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con3:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con3:Quan2:Item4', 'Extr_Num']      <- 36
		dat[dat$crossed=='Con4:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con4:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con4:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con4:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con4:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con4:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con4:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con4:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con4:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con4:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con4:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con4:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con4:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con4:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con4:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con4:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con4:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con4:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con4:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con4:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con4:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con4:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con4:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con4:Quan2:Item4', 'Extr_Num']      <- 36
		dat$crossed <- NULL
		# what side LEFT, MIDDLE, RIGHT corresponds with Expected, Borderline, Extreme?
		for (row in 1:nrow(dat)) {
			if (dat[row, 'Exp_Num']   == dat[row, 'Left'])  {dat[row, 'Exp_side']   <- 'left'}
			if (dat[row, 'Exp_Num']   == dat[row, 'Mid'])   {dat[row, 'Exp_side']   <- 'mid'}
			if (dat[row, 'Exp_Num']   == dat[row, 'Right']) {dat[row, 'Exp_side']   <- 'right'}
			if (dat[row, 'Bline_Num'] == dat[row, 'Left'])  {dat[row, 'Bline_side'] <- 'left'}
			if (dat[row, 'Bline_Num'] == dat[row, 'Mid'])   {dat[row, 'Bline_side'] <- 'mid'}
			if (dat[row, 'Bline_Num'] == dat[row, 'Right']) {dat[row, 'Bline_side'] <- 'right'}
			if (dat[row, 'Extr_Num']  == dat[row, 'Left'])  {dat[row, 'Extr_side']  <- 'left'}
			if (dat[row, 'Extr_Num']  == dat[row, 'Mid'])   {dat[row, 'Extr_side']  <- 'mid'}
			if (dat[row, 'Extr_Num']  == dat[row, 'Right']) {dat[row, 'Extr_side']  <- 'right'}
		}
		# what button press did the subject actually make? LEFT, MIDDLE, RIGHT, NOANSWER?
		dat$RESPONSE <- as.factor(dat$RESPONSE)
		# what number of dots corresponds with the subject's button press?
		for (row in 1 : nrow(dat) ) {
			switch(as.character(dat[row,'RESPONSE']),
						 'LEFT' = {dat[row, 'response_num'] <- dat[row, 'Left']},
						 'MIDDLE' = {dat[row, 'response_num'] <- dat[row, 'Mid']},
						 'RIGHT' = {dat[row, 'response_num'] <- dat[row, 'Right']},
						 'NOANSWER' = {dat[row, 'response_num'] <- NA}
			)
		}
		# what side was the subject's button-press? Left, mid right?
		dat$response_side <- tolower(dat$RESPONSE)
		dat$response_side[dat$response_side == "middle"] <- 'mid'
		dat$response_side <- factor(dat$response_side, exclude="noanswer")
		# what category was the subject's response? Expected, Borderline, Extreme
		dat$response_category <- "nocat"
		for (row in row.names(na.omit(dat))) {
			if (dat[row, 'response_num'] == dat[row, 'Exp_Num']) {dat[row, 'response_category'] <- 'expected'}
			if (dat[row, 'response_num'] == dat[row, 'Bline_Num']) {dat[row, 'response_category'] <- 'borderline'}
			if (dat[row, 'response_num'] == dat[row, 'Extr_Num']) {dat[row, 'response_category'] <- 'extreme'}
		}
		dat$response_category <- factor(dat$response_category, exclude="nocat")
		dat$RESPONSE <- NULL
		return(dat)
	} # end of classifyResponse function

	# declare local variables
	number_of_valid_subjects <- 30 # = 30
	number_of_rows <- 7680 # 7680
	number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
	dat <- gatherData(number_of_valid_subjects) # = 30
	dat <- classifyResponse(dat) # classify the response as expected, near, or far

	# SUBJECT
	dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
	# TRIAL
	dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)
	# make a centred Trial for modeling
	dat$c_Trl <-dat$Trial - mean(dat$Trial)
	# make a scaled Trial for modelling
	dat$s_Trl <- as.numeric(scale(dat$Trial))
	# ID
	# id is a unique identifier for the 7680 row data
	dat$id <- factor(paste(paste(dat$Subject),
												 paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
	# ITEM
	# create a centred numeric item variable for modeling
	dat$c_Itm <- ifelse(dat$Item==1, -.75,
											ifelse(dat$Item==2, -.25,
														 ifelse(dat$Item==3, .25, .75)))
	# make Item be a factor and assign labels
	dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
	# add ratios for Item
	item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
	# 0.2500000 0.4705882 0.5909091 0.6666667
	item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
	# -1.3441995 -0.1316642  0.5297187  0.9461450
	item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
	# 0.5125000 0.6876471 0.7691558 0.8166667
	item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))))
	# -1.37582241 -0.06614191  0.54334858 0.89861574
	dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
	dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
	dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
	dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
	dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
	dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
	dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187
	dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
	dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000
	dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471
	dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
	dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
	dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241
	dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191
	dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858
	dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
	# VAGUENESS
	# Create a factor coding for Vagueness
	dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
	dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
	dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
	dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
	dat$Vagueness <- as.factor(dat$Vagueness)
	# manually center Vagueness
	dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
	# NUMBER
	# Create a factor coding for Number use
	dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
	dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
	dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
	dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
	dat$Number <- as.factor(dat$Number)
	# manually center Number
	dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
	# CONDITION
	# make a factor out of Condition, as f_Cnd
	dat$f_Cnd <- factor(dat$Condition, levels=c(1,2,3,4), labels=c('Vg:Nm', 'Cr:Nm', 'Vg:Vb', 'Cr:Vb'))
	# ORDER
	# give the levels of Order meaningful names
	dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
	# make a manually centred Order
	dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
  # QUANTITY
	# give the levels of Quantity meaningful names
	dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
	# make a manually centred Quantity
	dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
	# INSTRUCTION
	# add number of characters in the instruction # 29 30 34 36 38
	dat$nchar_instr = nchar(dat$Instruction)
	dat$nchar_instr_scaled = as.vector(scale(nchar(dat$Instruction), scale=TRUE))
	# make Instruction be a factor (17 levels)
	dat$Instruction <- as.factor(dat$Instruction)
	# RT
	# add transformations of RT
	dat$RT_RecSqd    <- 1/dat$RT^2
	dat$RT_Rec       <- 1/dat$RT
	dat$RT_RecSqt    <- 1/sqrt(dat$RT)
	dat$RT_log       <- log(dat$RT)
	dat$RT_sqt       <- sqrt(dat$RT)
	dat$RT_raw       <- dat$RT
	dat$RT_sqd       <- dat$RT^2
  # print to file a table with information about the design
	design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order,
																						 Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
	design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
	row.names(design_info) <- NULL
	capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F),
								 file="design_info.table")
	# put dat in better column order
	dat <- subset(dat,
								select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
													 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
													 c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
													 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
													 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
													 Instruction, nchar_instr
								))
	# save dat as dat.Rda
	# This data set contains *all* trials 7680 including impossible trials and is
	# mainly for graphs comparing different removals
	save(dat, file="dat.Rda")
	# remove impossible trials from dat in new data dd #
	# Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998
	# i.e., throw out sticky fingers and timeouts,
	# and the trials that followed sticky fingers and timeouts
	# since they were likely affected by unusual previous trials.
	# lose impossible trials
	dd <- dat
	dd$RT[dd$RT == 1 ] <- NA
	dd$RT[dd$RT == 59998 ] <- NA
	dd <- dd[complete.cases(dd),]
	row.names(dd) <- NULL
	# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
	dd$RTprev <- NA
	for (s in levels(dd$Subject)) {
		nrows = nrow(dd[dd$Subject==s,])
		for (i in 1:nrows) {
			if (i==1) {
				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
			}
			else
				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
		}
	}
	# add transformations of previous RT
	dd$RTprev_RecSqd    <- 1/dd$RTprev^2
	dd$RTprev_Rec       <- 1/dd$RTprev
	dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev)
	dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
	dd$RTprev_log       <- log(dd$RTprev)
	dd$RTprev_sqt       <- sqrt(dd$RTprev)
	dd$RTprev_raw       <- dd$RTprev
	dd$RTprev_sqd       <- dd$RTprev^2
	# number of higher-level cells in the design (abstracting over Order) =
	# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
	# number of measurements in eaach cell = 8
	# 332 * 8 = 256
	# 256 * 30 = 7680
	dd$measurement=0
	dd$cell=0
	for (s in levels(dd$Subject)) {
		cellcount=0
		for (i in levels(dd$Item)) {
			for (n in levels(dd$Number)) {
				for (v in levels(dd$Vagueness)) {
					for (q in levels(dd$Quantity)) {
						measurementcount=0
						cellcount=cellcount+1
						for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
							measurementcount=measurementcount+1
							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
						}
					}
				}
			}
		}
	}
	# put dd in better column order
	dd <- subset(dd,
							 select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
							 					 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
							 					 c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
							 					 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
							 					 RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
							 					 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
							 					 Instruction, nchar_instr,
							 					 cell, measurement
							 ))
	save(dd, file="dd.Rda")
} # end of the else clause, which is run if both of the files dat.Rda and dd.Rda do not exist.
```

Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.

```{r plotDists, fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data
# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
dat_transforms <- dat
# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"
# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"
# create extra rows for dfs with removed (outlier) data points
dat1 <- dat_transforms                                       # nrow 7680
dat2 <- subset(dat_transforms, RT>1)                         # nrow 7678
dat2$dfSubset <- "RT>1"
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"
# bring the dfs back into main df
dat_transforms <- rbind(dat1, dat2, dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT>1", "RT>1&RT<59999", "RT>1&RT<30001"))
# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd")] <- NA
# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqd"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_Rec"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_Rec"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqd"]))
} # end of create normally distributed equivalents for the data transforms.
temp <- reshape2::melt(dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
             variable.name="transformation",
             value.name="score")
ggplot(temp, aes(score, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(dfSubset~transformation, scales="free", nrow=4) +
  theme_bw() + 
	theme(aspect.ratio=1) + 
	scale_color_manual(values=c("blue","red"))
```

# Now working with a data set without impossible trials.

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r boxPlots, fig.width=20, fig.height=5}
# allow reference to transformation
subdata = reshape2::melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

Show how mean times for individual subjects and items vary with respect to the grand mean RT.

```{r fastSlowRT, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), 
					 summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT),
					 ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0, "slow", "fast"))
itms=ddply(dd, .(level=Item), 
					 summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), 
					 ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0, "slow", "fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))
my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), 
              																	superpose.line=list(col = c("blue","red")))),
              groups=si$lsign,
              xlab=NULL,
              ylab="RT (in ms)",
              key=list(text=list(levels(si$lsign)), 
              				 lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, 
              				 					 pch=c(19,19), type='o'), divide=1),
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=list(15.5, 32.5), 
              									 labels=list("subject number", "item identifier"), cex=1 ),
                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), 
                          			 labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, 
                						 code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), 
                					 as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), 
                					 as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              }) # end of my1
my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from mean RT (in ms)",
              ylim=extendrange(si$ldiff,f=0.10)
							) # creates a second y-axis scale
print(doubleYScale(my1, my2, 
									 add.axis=TRUE, 
									 add.ylab2=TRUE, 
									 use.style=FALSE))
```

Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.

```{r fastSlowRecSqt, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.
subs <- ddply(dd, .(level=Subject), summarise, 
              effect="subject", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt), 
							ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt),
              lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
itms <- ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt),
              ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt), 
							lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
si <- rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))
my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings = list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups = si$lsign,
              xlab = NULL,
              ylab = "RT_RecSqt: reciprocal of square root of RT in ms\n",
              key=list(text=list(levels(si$lsign)), 
              				 lines=list(lty=c(1,1), 
              				 					 col=trellis.par.get('superpose.line')$col,
              				 					 pch=c(19,19), type='o'),
              				 divide=1),
              mean = unique(si$emean),
              scales = list(x=list(relation="free", at=list(15.5, 32.5), 
              										 labels=list("subject number", "item identifier"), cex=1 ) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, 
                						 groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), 
                					 as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), 
                					 as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })
my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="\nDeviation from mean RT_RecSqt (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))
print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

Show how practice and fatigue effects vary over subjects

```{r pracFatigueSubjs, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + 
	facet_wrap(~Subject,nrow=3) + 
	theme(panel.grid=element_blank(), 
				panel.margin = unit(0, "lines"), 
				panel.background = element_rect(fill = 'white', colour='black' ), 
				legend.position="top") + 
	scale_x_continuous("Trial", breaks=NULL) + 
	geom_point(pch=19, cex=.25, alpha=.5) + 
	stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + 
	theme(aspect.ratio = 1) 
```

Show practice and fatigue effects vary over items.

```{r pracFatigueItms, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + 
	facet_grid(~Item) + 
	theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), 
				panel.background = element_rect(fill = 'white', colour='black' ), 
				legend.position="top") + 
	scale_x_continuous("Trial") + 
	geom_point(pch=19, cex=.25, alpha=.5) + 
	stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + 
	theme(aspect.ratio = 1) 
```

# Plot main effects and interactions

Main effects

Plot main effects on several transformations

```{r plotMainsa, fig.width=16, fig.height=6, warning=FALSE}
# plot main effects on observed data
# allow reference to transformation
temp <- reshape2::melt(dd,
             id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
             variable.name="transformation",
             value.name="score")
# allow reference to effect
temp2 <- reshape2::melt(temp, 
              id.vars=c("transformation", "score"),
              measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
              variable.name="effect",
              value.name="level")
# do the plot
ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
  facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
  stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
  stat_summary(fun.y=mean, geom="line") +
  theme(aspect.ratio=1, axis.text.y=element_blank())
```

Main effects

```{r mainB, fig.width=9, fig.height=3}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
# plot main effects
ggplot(rts, aes(y=RT_RecSqt, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r mainD, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Number")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
ggplot(rts, 
			 aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, 
			 		ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
  scale_color_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
```

2 way interactions not involving items.

```{r 2way, fig.width=10, fig.height=3}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Number")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Number",4)
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Number:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Number:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, RT_RecSqt_norm, sd, se, ci))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
# plot 2 way interactions
ggplot(rts, aes(y=RT_RecSqt, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

```{r 22way, message=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Number"
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Number"
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Number"
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"
rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
rts$F3 <- factor(paste(rts$F1,rts$F2))
rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + 
	geom_line() + 
	theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
```

Plot vagueness by number interaction over items

```{r plotInt1, fig.width=7, fig.height=3.5, eval=TRUE, cache=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=condition, shape=condition, fill=condition)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(size=4, position=dodge) + 
  scale_shape_manual("",values = c(22, 22, 21, 21)) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
  			legend.key = element_blank(), aspect.ratio=1, 
  			axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number)
```

3 way interactions

```{r plot3way, fig.width=10,fig.height=5}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p1=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Quantity, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p2=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p3=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p4=ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
grid.arrange(nrow=2, p1,p2,p3,p4)
```

Plot vagueness by number by quantity over items.

```{r vnitems, fig.width=10, fig.height=4, eval=TRUE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness, 
                fill=Vagueness)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(pch=21, size=4, position=dodge) + 
  scale_fill_manual("",values=c("black","white")) +
  ggtitle("Response time") + 
  ylab("RT_Rec_Sqt") + 
  xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background=element_rect(fill="white",color=1),
        legend.key = element_blank(), 
        aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number+Quantity)
```

4 way interaction

```{r 4way}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, group=Quantity, pch=Quantity)) +
  facet_grid(~Order+Number, scale="free_x") +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank())
```

4 way interaction split over items

```{r 4waysplit, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  geom_line() + 
  geom_errorbar(width=.12) + 
  geom_point(pch=21, size=2) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

4 way interaction split over items showing ratio of item dots.

```{r 4waySplitRatio, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
#dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  # geom_line(position=dodge) + 
	geom_line() +
  geom_errorbar(width=.02) + 
  geom_point(pch=21, size=2) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), 
  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

# Show compression of item ratios

```{r showCompression, fig.width=10, fig.height=7, echo=F}
#  Show compression of ratios of numbers of dots in the items
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)
plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')
points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
text(y=1.15, x=0.2, labels="mean (ratio min/med,\n            ratio med/max)", col='red')
points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
axis(1, at=seq(0,1,0.1))
```

## ols model v4, RecSqt, in progress

Pairs plot

```{r pairs, fig.width=7, fig.height=7}
pairs(dd[,c("RT_RecSqt", "Number", "Vagueness", "RTprev_RecSqt")], pch = 19, cex=.075)
```

Build the model

```{r doModelols}
## do ols model
# define datadist
dd.dd = datadist(dd)
options(datadist = "dd.dd")
# build model
v4 <- ols(data=dd, 
          RT_RecSqt ~ 
            c_Vag + c_Num + c_Qty + c_Ord + 
            c_Num:c_Vag:c_Qty+
            item_mean_ratio + 
            #c_Vag:c_Num + 
            #c_Vag:c_Num:item_mean_ratio + 
            # pol(s_Trl, 3) +
            s_Trl+
            # pol(RTprev_RecSqt,3) +
            RTprev_RecSqt +
            nchar_instr
            # cell+
            # measurement,
            ,
          x=T, y=T )
```

```{r likRat}
# print out p value for likelihood ratio with df
cat("lik.ratio p value, whether the model as a whole is explanatory: p =", 1-pchisq(v4$stats[2],v4$stats[3]))
# print out R2 for the model
cat("R^2 = ", v4$stats[4])
```

Show the model tables

```{r modTabs}
# show model
v4
# show model summary
summary(v4)
# do anova 
an1 = anova(v4)
# show anova
an1
```

plot 'partial effects'

```{r partial, fig.width=12, fig.height=5}
## plot partial effects
# Compute Predicted Values and Confidence Limits
p1=Predict(v4)
# plot 'partial effects'
plot(p1, anova=an1, pval=T, aspect=1, main="ols model v4")
```

```{r daz99}
# add predicted RT_RecSqt to dd
dd$RT_Predicted <- predict (v4)
```

Plot model coefficients and ci's

```{r coefsAndCiS}
par(las=2, mar=c(14,6,1,1))
y=coef(v4) # with intercept
n=length(y)
y0=confint(v4)[,1]
y1=confint(v4)[,2]
y=coef(v4)[-1] # omit intercept
n=length(y)
y0=confint(v4)[-1,1]
y1=confint(v4)[-1,2]
plot(y, xaxt="n", xlab="", ylab="RT_RecSqt\n", pch=19, ylim=extendrange(y, f=.10), main="Speed")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

Collinearity is assessed by the condition number $k$.\
The greater the collinearity, the closer the matrix of predictors is is to becoming _singular_.\
$k$ estimates the extent to which a matrix is $\text{singular}$\
$0\ldots6$ no collinearity\
$\text{around} 15$ medium collinearity\
$>30$ indicates potentially harmful collinearity\

```{r kis}
cat("k is:",
collin.fnc(dd[,	c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", 
									"RTprev_RecSqt", "nchar_instr", "measurement")
							])$cnumber
)
```

```{r plotg, fig.width=6, fig.height=5}
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(
	as.matrix(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord",
									"item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])))
```

Model criticism baayen plots

```{r baayenMplots, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism
par(mfrow=c(1,4), pty='s')
# create scaled residuals
v4$rstand = as.vector(scale(resid(v4)))
# plot scaled residuals density
plot(density(v4$rstand), main='density of\nscaled residuals')
# plot sample quantiles versus theoretical quantiles
qqnorm(v4$rstand, cex=.5)
qqline(v4$rstand)
# plot standardised residuals versus fitted values
plot(v4$rstand ~ fitted(v4), pch='.', main='std resid\nv fitted')
# absolute standardised residuals greater than 2.5 are candidates 
# for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v4, "dffits"))
# plot dffits
plot(dffits, type='h', main='dffits')
# here no overly influential
w = which.influence(fit = v4, cutoff = 0.4)
if (length(w)==0) {
	print('There are no overly influential data points')
} else {
	print('There are some overly influential variables:');
	print(w)
}
```

Model validation

Validation tests overfitting using bootstrap.

Bootstrapping draws as a bootstrap sample the same number of samples as the original data, at random with replacement, from the original data.

Then fit the model to the data in the bootstrap sample, and use this model to predict the reaction times for the original full data (which contains many samples that are new to the bootstrap model).

Then compare the goodness of fit of the bootstrap model with the goodness of fit of the original model.

Averaged over a large number of bootstrap models these comparisons of goodness of fit  reveal to what extent the original model overfits the data.

Function validate() in rms does this _re-sampling validation_.

```{r val_1, echo=TRUE}
# argument B is the number of bootstrap runs
# argument pr is whether to print the results of each run
v_1 <- validate(v4, bw = T, B = 200, pr=FALSE, estimates=TRUE)
```

```{r prval}
# in print.validate, B is the number of re-samples to show outocmes for each resample
print(v_1, B=0)
```

```{r kept}
# show how often each number of variables kept was observed across the 200 runs
xtabs(~rowSums(attr(v_1,"kept")))
```

# model lmer

```{r rmme}
load("dd.Rda")
```

```{r lmer5}
v5 <- lmerTest::lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + c_Num + c_Qty + c_Ord + 
             c_Num:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject)
)
```

```{r v51}
summary(v5)
```

```{r v52}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r v53}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

Plot model coefficients and ci's

```{r v54}
e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
q=as.data.frame(confint(v5, method='Wald')[18:26,1:2])
eq=cbind(e,q)
y=eq[rev(rownames(eq)),]
par(mar=c(4,10,1,1))
plot(y=1:nrow(y), x=y$coef, xlim=range(y$"2.5 %", y$"97.5 %"), type='n', axes=F, xlab="", ylab="")
segments(y0=1:nrow(y), y1=1:nrow(y), x0=y$"2.5 %", x1=y$"97.5 %", lwd=.55)
points(y=1:nrow(y), x=y$coef, pch=20, cex=.75)
abline(v=0, lty=3, col='grey')
axis(2, labels=row.names(y), at=1:nrow(y), las=1)
axis(1, las=1)
mtext("estimate", side=1, line=2.5, cex.lab=1, las=1)
box(col='grey')
```

Model criticism baayen plots

```{r baayenPLots99, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism
par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
```