### 2016-08-07 experiments/e2/writeup/e2.Rmd
tidied up e2 considerably, switched to pdf output

---
title: "Experiment 2 (of 4)"
author: "Matt Green"
output: 
  pdf_document:
    toc: true
---

```{r loadKnitr, echo=FALSE}
suppressPackageStartupMessages(library(knitr))
```

```{r loadLibs, echo=FALSE, message=FALSE}
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
```

```{r get_the_data, echo=FALSE}
# if the file data.Rda already exists then load it, else do data wrangling
if( file.exists('data_processed.Rda') ) {
	load('data_processed.Rda')
	} else {
	gatherData = function(number_of_valid_subjects) {
		data_dir <- '../experimentCode/output/'
		column.headers.df <-  head( read.table(
			paste(data_dir,'subject01.data',sep=''),
			header=TRUE),0)
		gathered.data <- column.headers.df
		for (subject in 1:number_of_valid_subjects) {
			current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
			current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
			gathered.data <- rbind(gathered.data, current.data)
		}
		return(gathered.data)
	} # end of gatherData function
	classifyResponse = function(dat) {
		# what were they expected to respond?
		dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', dat$Quantity, ':Item', dat$Item, sep=''))
		dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con1:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con1:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con1:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con1:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con1:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con1:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con1:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con1:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con1:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con1:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con1:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con1:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con1:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con1:Quan2:Item4', 'Extr_Num']      <- 36
		dat[dat$crossed=='Con2:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con2:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con2:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con2:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con2:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con2:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con2:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con2:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con2:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con2:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con2:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con2:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con2:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con2:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con2:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con2:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con2:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con2:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con2:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con2:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con2:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con2:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con2:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con2:Quan2:Item4', 'Extr_Num']      <- 36
		dat[dat$crossed=='Con3:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con3:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con3:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con3:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con3:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con3:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con3:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con3:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con3:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con3:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con3:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con3:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con3:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con3:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con3:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con3:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con3:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con3:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con3:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con3:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con3:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con3:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con3:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con3:Quan2:Item4', 'Extr_Num']      <- 36
		dat[dat$crossed=='Con4:Quan1:Item1', 'Exp_Num'] <-  6
		dat[dat$crossed=='Con4:Quan1:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con4:Quan1:Item1', 'Extr_Num']      <- 24
		dat[dat$crossed=='Con4:Quan1:Item2', 'Exp_Num'] <- 16
		dat[dat$crossed=='Con4:Quan1:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con4:Quan1:Item2', 'Extr_Num']      <- 34
		dat[dat$crossed=='Con4:Quan1:Item3', 'Exp_Num'] <- 26
		dat[dat$crossed=='Con4:Quan1:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con4:Quan1:Item3', 'Extr_Num']      <- 44
		dat[dat$crossed=='Con4:Quan1:Item4', 'Exp_Num'] <- 36
		dat[dat$crossed=='Con4:Quan1:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con4:Quan1:Item4', 'Extr_Num']      <- 54
		dat[dat$crossed=='Con4:Quan2:Item1', 'Exp_Num'] <- 24
		dat[dat$crossed=='Con4:Quan2:Item1', 'Bline_Num']     <- 15
		dat[dat$crossed=='Con4:Quan2:Item1', 'Extr_Num']      <-  6
		dat[dat$crossed=='Con4:Quan2:Item2', 'Exp_Num'] <- 34
		dat[dat$crossed=='Con4:Quan2:Item2', 'Bline_Num']     <- 25
		dat[dat$crossed=='Con4:Quan2:Item2', 'Extr_Num']      <- 16
		dat[dat$crossed=='Con4:Quan2:Item3', 'Exp_Num'] <- 44
		dat[dat$crossed=='Con4:Quan2:Item3', 'Bline_Num']     <- 35
		dat[dat$crossed=='Con4:Quan2:Item3', 'Extr_Num']      <- 26
		dat[dat$crossed=='Con4:Quan2:Item4', 'Exp_Num'] <- 54
		dat[dat$crossed=='Con4:Quan2:Item4', 'Bline_Num']     <- 45
		dat[dat$crossed=='Con4:Quan2:Item4', 'Extr_Num']      <- 36
		dat$crossed <- NULL
		# what side LEFT, MIDDLE, RIGHT corresponds with Expected, Borderline, Extreme?
		for (row in 1:nrow(dat)) {
			if (dat[row, 'Exp_Num']   == dat[row, 'Left'])  {dat[row, 'Exp_side']   <- 'left'}
			if (dat[row, 'Exp_Num']   == dat[row, 'Mid'])   {dat[row, 'Exp_side']   <- 'mid'}
			if (dat[row, 'Exp_Num']   == dat[row, 'Right']) {dat[row, 'Exp_side']   <- 'right'}
			if (dat[row, 'Bline_Num'] == dat[row, 'Left'])  {dat[row, 'Bline_side'] <- 'left'}
			if (dat[row, 'Bline_Num'] == dat[row, 'Mid'])   {dat[row, 'Bline_side'] <- 'mid'}
			if (dat[row, 'Bline_Num'] == dat[row, 'Right']) {dat[row, 'Bline_side'] <- 'right'}
			if (dat[row, 'Extr_Num']  == dat[row, 'Left'])  {dat[row, 'Extr_side']  <- 'left'}
			if (dat[row, 'Extr_Num']  == dat[row, 'Mid'])   {dat[row, 'Extr_side']  <- 'mid'}
			if (dat[row, 'Extr_Num']  == dat[row, 'Right']) {dat[row, 'Extr_side']  <- 'right'}
		}
		# what button press did the subject actually make? LEFT, MIDDLE, RIGHT, NOANSWER?
		dat$RESPONSE <- as.factor(dat$RESPONSE)
		# what number of dots corresponds with the subject's button press?
		for (row in 1 : nrow(dat) ) {
			switch(as.character(dat[row,'RESPONSE']),
						 'LEFT' = {dat[row, 'response_num'] <- dat[row, 'Left']},
						 'MIDDLE' = {dat[row, 'response_num'] <- dat[row, 'Mid']},
						 'RIGHT' = {dat[row, 'response_num'] <- dat[row, 'Right']},
						 'NOANSWER' = {dat[row, 'response_num'] <- NA}
			)
		}
		# what side was the subject's button-press? Left, mid right?
		dat$response_side <- tolower(dat$RESPONSE)
		dat$response_side[dat$response_side == "middle"] <- 'mid'
		dat$response_side <- factor(dat$response_side, exclude="noanswer")
		# what category was the subject's response? Expected, Borderline, Extreme
		dat$response_category <- "nocat"
		for (row in row.names(na.omit(dat))) {
			if (dat[row, 'response_num'] == dat[row, 'Exp_Num']) {dat[row, 'response_category'] <- 'expected'}
			if (dat[row, 'response_num'] == dat[row, 'Bline_Num']) {dat[row, 'response_category'] <- 'borderline'}
			if (dat[row, 'response_num'] == dat[row, 'Extr_Num']) {dat[row, 'response_category'] <- 'extreme'}
		}
		dat$response_category <- factor(dat$response_category, exclude="nocat")
		dat$RESPONSE <- NULL
		return(dat)
	} # end of classifyResponse function

	# declare local variables
	number_of_valid_subjects <- 30 # = 30
	number_of_rows <- 7680 # 7680
	number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
	dat <- gatherData(number_of_valid_subjects) # = 30
	dat <- classifyResponse(dat) # classify the response as expected, near, or far

	# SUBJECT
	dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
	# TRIAL
	dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)
	# make a centred Trial for modeling
	dat$c_Trl <-dat$Trial - mean(dat$Trial)
	# make a scaled Trial for modelling
	dat$s_Trl <- as.numeric(scale(dat$Trial))
	# ID
	# id is a unique identifier for the 7680 row data
	dat$id <- factor(paste(paste(dat$Subject),
												 paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
	# ITEM
	# create a centred numeric item variable for modeling
	dat$c_Itm <- ifelse(dat$Item==1, -.75,
											ifelse(dat$Item==2, -.25,
														 ifelse(dat$Item==3, .25, .75)))
	# make Item be a factor and assign labels
	dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
	# add ratios for Item
	item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
	# 0.2500000 0.4705882 0.5909091 0.6666667
	item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
	# -1.3441995 -0.1316642  0.5297187  0.9461450
	item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
	# 0.5125000 0.6876471 0.7691558 0.8166667
	item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))))
	# -1.37582241 -0.06614191  0.54334858 0.89861574
	dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
	dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
	dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
	dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
	dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
	dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
	dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187
	dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
	dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000
	dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471
	dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
	dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
	dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241
	dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191
	dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858
	dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
	# VAGUENESS
	# Create a factor coding for Vagueness
	dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
	dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
	dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
	dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
	dat$Vagueness <- as.factor(dat$Vagueness)
	# manually center Vagueness
	dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
	# NUMBER
	# Create a factor coding for Number use
	dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
	dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
	dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
	dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
	dat$Number <- as.factor(dat$Number)
	# manually center Number
	dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
	# CONDITION
	# make a factor out of Condition, as f_Cnd
	dat$f_Cnd <- factor(dat$Condition, levels=c(1,2,3,4), labels=c('Vg:Nm', 'Cr:Nm', 'Vg:Vb', 'Cr:Vb'))
	# ORDER
	# give the levels of Order meaningful names
	dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
	# make a manually centred Order
	dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
  # QUANTITY
	# give the levels of Quantity meaningful names
	dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
	# make a manually centred Quantity
	dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
	# INSTRUCTION
	# add number of characters in the instruction # 29 30 34 36 38
	dat$nchar_instr = nchar(dat$Instruction)
	dat$nchar_instr_scaled = as.vector(scale(nchar(dat$Instruction), scale=TRUE))
	# make Instruction be a factor (17 levels)
	dat$Instruction <- as.factor(dat$Instruction)
	# RT
	# add transformations of RT
	dat$RT_RecSqd    <- 1/dat$RT^2
	dat$RT_Rec       <- 1/dat$RT
	dat$RT_RecSqt    <- 1/sqrt(dat$RT)
	dat$RT_log       <- log(dat$RT)
	dat$RT_sqt       <- sqrt(dat$RT)
	dat$RT_raw       <- dat$RT
	dat$RT_sqd       <- dat$RT^2
  # print to file a table with information about the design
	design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order,
																						 Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
	design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
	row.names(design_info) <- NULL
	capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F),
								 file="design_info.table")
	# put dat in better column order
	dat <- subset(dat,
								select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
													 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
													 c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
													 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
													 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
													 Instruction, nchar_instr
								))
	# This data set (dat) contains *all* trials 7680 including impossible trials and is
	# mainly for graphs comparing different removals
	save(dat, file='data_raw.Rda')
	# remove impossible trials from dat in new data dd #
	# Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998
	# i.e., throw out sticky fingers and timeouts,
	# and the trials that followed sticky fingers and timeouts
	# since they were likely affected by unusual previous trials.
	# lose impossible trials
	dd <- dat
	dd$RT[dd$RT == 1 ] <- NA
	dd$RT[dd$RT == 59998 ] <- NA
	dd <- dd[complete.cases(dd),]
	row.names(dd) <- NULL
	# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
	dd$RTprev <- NA
	for (s in levels(dd$Subject)) {
		nrows = nrow(dd[dd$Subject==s,])
		for (i in 1:nrows) {
			if (i==1) {
				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
			}
			else
				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
		}
	}
	# add transformations of previous RT
	dd$RTprev_RecSqd    <- 1/dd$RTprev^2
	dd$RTprev_Rec       <- 1/dd$RTprev
	dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev)
	dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
	dd$RTprev_log       <- log(dd$RTprev)
	dd$RTprev_sqt       <- sqrt(dd$RTprev)
	dd$RTprev_raw       <- dd$RTprev
	dd$RTprev_sqd       <- dd$RTprev^2
	# number of higher-level cells in the design (abstracting over Order) =
	# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
	# number of measurements in eaach cell = 8
	# 332 * 8 = 256
	# 256 * 30 = 7680
	dd$measurement=0
	dd$cell=0
	for (s in levels(dd$Subject)) {
		cellcount=0
		for (i in levels(dd$Item)) {
			for (n in levels(dd$Number)) {
				for (v in levels(dd$Vagueness)) {
					for (q in levels(dd$Quantity)) {
						measurementcount=0
						cellcount=cellcount+1
						for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
							measurementcount=measurementcount+1
							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
						}
					}
				}
			}
		}
	}
	# put dd in better column order
	dd <- subset(dd,
							 select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
							 					 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
							 					 c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
							 					 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
							 					 RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
							 					 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
							 					 Instruction, nchar_instr,
							 					 cell, measurement
							 ))
	save(dd, file="data_processed.Rda")
} # end of the else clause, which is run if the file data.Rda does not exist.
```

Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.

```{r plotDists, fig.width=5, fig.height=5, message=FALSE, echo=FALSE}
# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
load('data_raw.Rda') # dat
dat_transforms <- dat
# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"
# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"
# create extra rows for dfs with removed (outlier) data points
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"
# bring the dfs back into main df
dat_transforms <- rbind(dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("RT>1&RT<59999", "RT>1&RT<30001"))
# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_log", "RT_raw")] <- NA
# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  }# end of create normally distributed equivalents for the data transforms.
temp <- reshape2::melt(dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_log", "RT_raw"),
             variable.name="transformation",
             value.name="time")
ggplot(temp, aes(time, colour=RTtype)) + geom_freqpoly() + facet_wrap(transformation~dfSubset, scales='free') + scale_color_manual(values=c("blue","red")) + theme_bw() + theme(aspect.ratio=1)
```

# Now working with a data set without impossible trials.

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r boxPlots, fig.width=7, fig.height=3, echo=FALSE}
load('data_processed.Rda')
# allow reference to transformation
subdata = reshape2::melt(dd,
               measure.vars=c("RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
  xlab("") + ylab("")
```

Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.

```{r fastSlowRT, fig.width=7, fig.height=3, message=FALSE, echo=FALSE}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), 
					 summarise, effect="subject", emean=mean(dd$RT_log), lmean=mean(RT_log),
					 ldiff=mean(RT_log)-mean(dd$RT_log))
itms=ddply(dd, .(level=Item), 
					 summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
					 ldiff=mean(RT_log)-mean(dd$RT_log))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col='black'), 
              																	superpose.line=list(col='black'))),
              xlab=NULL,
              ylab=NULL,
              main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=NULL, 
              									 labels=list("subject number", "item identifier"), cex=1 ),
                          y=list(at=    c(0,7,unique(si$emean),8,9),
                          			 labels=c("0","7","Mean Log RT","8","9"))),
              panel=function(x,y,type,subscripts,...){
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, subscripts=subscripts, code=3, length=0, ...)
                panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
              }) # end of my1
my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from\nmean Log RT",
              ylim=extendrange(si$ldiff,f=0.10)
							) # creates a second y-axis scale
print(doubleYScale(my1, my2, 
									 add.axis=TRUE, 
									 add.ylab2=TRUE, 
									 use.style=FALSE))
```

# Plot main effects and interactions

Main effects

Plot main effects on several transformations

```{r plotMainsa, fig.width=6, fig.height=6, warning=FALSE, echo=FALSE}
# plot main effects on observed data
# allow reference to transformation
temp <- reshape2::melt(dd,
             id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
             variable.name="transformation",
             value.name="score")
# allow reference to effect
temp2 <- reshape2::melt(temp, 
              id.vars=c("transformation", "score"),
              measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
              variable.name="effect",
              value.name="level")
# do the plot
ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
  facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
  stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
  stat_summary(fun.y=mean, geom="line") +
  theme(aspect.ratio=1, axis.text.y=element_blank())
```

Main effects

```{r mainB, fig.width=6, fig.height=3, echo=FALSE}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
# plot main effects
ggplot(rts, aes(y=RT_RecSqt, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r mainD, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Number")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
ggplot(rts, 
			 aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, 
			 		ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
  scale_color_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
```

2 way interactions not involving items.

```{r 2way, fig.width=6, fig.height=3, echo=FALSE}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Number")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Number",4)
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Number:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Number:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, sd, se, ci))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
# plot 2 way interactions
ggplot(rts, aes(y=RT_RecSqt, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

```{r 22way, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Number"
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Number"
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Number"
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"
rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
rts$F3 <- factor(paste(rts$F1,rts$F2))
rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + 
	geom_line() + 
	theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
```

Plot vagueness by number interaction over items

```{r plotInt1, fig.width=7, fig.height=3.5, echo=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=condition, shape=condition, fill=condition)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(size=4, position=dodge) + 
  scale_shape_manual("",values = c(22, 22, 21, 21)) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
  			legend.key = element_blank(), aspect.ratio=1, 
  			axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number)
```

3 way interactions

```{r plot3way, fig.width=7, fig.height=5, echo=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p1=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Quantity, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p2=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p3=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p4=ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
grid.arrange(nrow=2, p1,p2,p3,p4)
```

Plot vagueness by number by quantity over items.

```{r vnitems, fig.width=7, fig.height=4, echo=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness, 
                fill=Vagueness)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(pch=21, size=4, position=dodge) + 
  scale_fill_manual("",values=c("black","white")) +
  ggtitle("Response time") + 
  ylab("RT_Rec_Sqt") + 
  xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background=element_rect(fill="white",color=1),
        legend.key = element_blank(), 
        aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number+Quantity)
```

4 way interaction

```{r 4way, fig.width=6, fig.height=3, echo=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, group=Quantity, pch=Quantity)) +
  facet_grid(~Order+Number, scale="free_x") +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank())
```

4 way interaction split over items

```{r 4waysplit, fig.width=7, fig.height=7, echo=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  geom_line() + 
  geom_errorbar(width=.12) + 
  geom_point(pch=21, size=2) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

4 way interaction split over items showing ratio of item dots.

```{r 4waySplitRatio, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
#dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  # geom_line(position=dodge) + 
	geom_line() +
  geom_errorbar(width=.02) + 
  geom_point(pch=21, size=2) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), 
  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

# Show compression of item ratios

```{r showCompression, fig.width=7, fig.height=7, echo=FALSE}
#  Show compression of ratios of numbers of dots in the items
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)
plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')
points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
text(y=1.15, x=0.2, labels="mean (ratio min/med,\n            ratio med/max)", col='red')
points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
axis(1, at=seq(0,1,0.1))
```

# Lmer model

```{r rmme}
load("data_processed.Rda")
```

```{r lmer5}
v5 <- lmerTest::lmer(data=dd, 
           RT_log ~ 
             c_Vag + c_Num + c_Qty + c_Ord + 
             c_Num:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_log +
             nchar_instr +
             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
v5b <- lme4::lmer(data=dd, 
           RT_log ~ 
             c_Vag + c_Num + c_Qty + c_Ord + 
             c_Num:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_log +
             nchar_instr +
             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{r v51}
summary(v5)
```

```{r v511, results='asis'}
print(xtable(coef(summary(v5b))))
```

```{r v52}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r v53}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

Plot model coefficients and ci's

```{r v54, fig.width=7, fig.height=4}
e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
q=as.data.frame(confint(v5, method='Wald')[18:26,1:2])
eq=cbind(e,q)
y=eq[rev(rownames(eq)),]
par(mar=c(4,10,1,1))
plot(y=1:nrow(y), x=y$coef, xlim=range(y$"2.5 %", y$"97.5 %"), type='n', axes=F, xlab="", ylab="")
segments(y0=1:nrow(y), y1=1:nrow(y), x0=y$"2.5 %", x1=y$"97.5 %", lwd=.55)
points(y=1:nrow(y), x=y$coef, pch=20, cex=.75)
abline(v=0, lty=3, col='grey')
axis(2, labels=row.names(y), at=1:nrow(y), las=1)
axis(1, las=1)
mtext("estimate", side=1, line=2.5, cex.lab=1, las=1)
box(col='grey')
```

Model criticism baayen plots

```{r baayenPLots99, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism
par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
```