### 2016-09-17 papers/2016-Springer-Chapter/v-chapter-tables.tex
new structure

\begin{table}
\caption{Table of instructions for the pair (5,25). Experiment 1}
\label{instructionse1} 
\begin{tabular}{lll}
\hline\noalign{\smallskip}
vagueness&example\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& 	Choose the square with 5 dots \\
vague	&	Choose the square with few dots\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
\label{instructionse2} 
\begin{tabular}{llll}
\hline\noalign{\smallskip}
vagueness&instruction format&selection task&example\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& 	numeric	& matching	&	Choose the square with 6 dots \\
vague	&	numeric 	& matching	&	Choose a square with about 10 dots\\
crisp		&	verbal	& comparison	&	Choose the square with the fewest dots\\
vague	&	verbal	& comparison	&	Choose a square with few dots\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Experiment 3 Instructions assuming 6,15,24 dots as the Item, and showing fewer instead of more}
\label{instructionse3} 
\begin{tabular}{llll}
\hline\noalign{\smallskip}
vagueness&instruction format&selection task&instruction\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& numeric		&matching 	& Choose a square with 6 dots \\ 
crisp 	& numeric		&comparison 	& Choose a square with fewer than 20 dots \\
vague 	& numeric		&matching 	& Choose a square with about 10 dots \\ 
vague 	& numeric		&comparison 	& Choose a square with far fewer than 20 dots \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Experiment 4: Instructions}
\label{instructionse4}
\begin{tabular}{llll}
\hline\noalign{\smallskip}
vagueness&instruction format& selection task&instruction\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& verbal	&matching 	& Choose a square with the same number of dots as the target \\ 
crisp 	& verbal	&comparison	& Choose a square with fewer dots than the target \\
vague 	& verbal	&matching 	& Choose a square with about the same number of dots as the target \\ 
vague 	& verbal	&comparison	& Choose a square with far fewer dots than the target \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Vagueness as range reduction}
\label{vaguenessasrangereduction} 
\begin{tabular}{lll}
\hline\noalign{\smallskip}
vagueness & selection task & candidates\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& matching 	& 1 candidate\\
crisp 	& comparison 	& 2 candidates\\
vague 	& matching 	& 2 candidates\\
vague 	& comparison 	& 1 candidate\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}