### 2016-09-25 experiments/e1/writeup/e1.Rmd
changed number of digits in some tables

---
title: "Experiment 1 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

```{r pander, echo=FALSE, cache=FALSE}
library(pander)
library(knitr)
```

```{r showInstructions, echo=FALSE, cache=TRUE}
v = c("crisp","vague")
e <- c("Choose the the square with many dots", "Choose the square with 20 dots")
i = data.frame(v, e)
names(i) = c("vagueness", "example instruction")
pander(i, justify=c('center', 'left'))
```

```{r showItems, echo=FALSE, cache=TRUE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Gap = c("05","10","15","20"),
               x = c("25:20","25:15","25:10","25:05"),
               z = c("25:30","25:35","25:40","25:45") )
names(i) = c('Gap Size', 'Small Target', 'Large Target')
pander(i, justify=c('center', 'center', 'center'))
```

```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R') 
```

In this experiment it was possible to identify errors.
This is because there were only 2 squares and an instruction uniquely identified one of them.
Therefore error trials were removed from analysis of response times:

```{r separateDataFramesForRTandErrorAnalyses, echo=TRUE, cache=FALSE}
rawdata <- read.table('data.txt', sep=' ')
rtdata <- subset(rawdata, error==FALSE)
erdata <- rawdata
```

Correct responses were trimmed at 2.5 sd separately for each subject:

```{r trimData, echo=TRUE, cache=FALSE}
tempdata <- perSubjectTrim.fnc(rtdata, response='rt', subject='subject', trim = 2.5)$data
rtdata <- tempdata
```

This code creates the condition means with confidence intervals used for plotting:

```{r computeSummariesForGraphs, cache=TRUE}
rtdataplot1 = summarySEwithin2(rtdata,  measurevar="logrt",  withinvars=c("Vgn","Gap"), idvar="subject")
erdataplot1 <- summarySEwithin2(erdata, measurevar="error", withinvars=c("Vgn","Gap"), idvar="subject")
```

```{r rtplot, echo = FALSE}
dodge=.06
rtplot = ggplot(rtdataplot1, aes(y=logrtNormed, x=Gap, group=Vgn, fill=Vgn, shape=Vgn, ymin=logrtNormed-ci, ymax=logrtNormed+ci)) + 
  geom_line(position=position_dodge(dodge)) + 
  geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
  geom_point(pch=21, size=2, position=position_dodge(dodge)) +
  ggtitle("Response time") + 
  ylab("Respone time (log (ms))") + 
  xlab("Gap Size") +
  scale_fill_grey(name=element_blank(), start=0, end=1, guide=FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() ) +
  theme(aspect.ratio=1) + xlab("")
```

```{r erplot, echo=FALSE}
dodge=.06
erplot = ggplot(erdataplot1, aes(y=errorNormed, x=Gap, group=Vgn, fill=Vgn, shape=Vgn, ymin=errorNormed-ci, ymax=errorNormed+ci)) + 
  geom_line(position=position_dodge(dodge)) + 
  geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
  geom_point(pch=21, size=2, position=position_dodge(dodge)) +
  ggtitle("Error rate") + 
  ylab("Error rate (proportion of errors)") + 
  xlab("Gap Size") +
  scale_fill_grey(name=element_blank(), start=0, end=1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() )+
  theme(aspect.ratio=1) + xlab("")
```

```{r plotRTandErrorRates, fig.width=7.5, fig.height=3, message=FALSE, echo=FALSE}
grid.arrange(rtplot, erplot, ncol=2)
```

This code computes the models:

```{r showVersions, echo=FALSE, comment=NA}
cat (paste(sep="", 
           c("lmerTest version "), c(packageDescription("lmerTest", fields = "Version")),
           "; ", 
           c("lme4 version "), c(packageDescription("lme4", fields = "Version"))) ) 
```

```{r computeLmerModels, cache = TRUE, tidy=FALSE}
rt.lmer.1 <- lmerTest::lmer(
  data=rtdata, logrt ~ c_Vgn * c_Gap + (c_Vgn * c_Gap|subject))

ac.lmer.1 <- lme4::glmer(
  data = erdata, error ~ c_Vgn * c_Gap + (c_Vgn * c_Gap | subject), 
  family="binomial", control = glmerControl(optimizer = "bobyqa"))
```

# Response times

```{r, results='asis', echo=F}
print(
  xtable(rtdataplot1,
         caption='Condition means'),
  include.rownames=FALSE,
  caption.placement='top',
  type='html')
```

```{r tableRTmodelSimple, results='asis', echo=FALSE, cache=TRUE}
x = as.data.frame(format(summary(rt.lmer.1)$coef, scientific=FALSE), stringsAsFactors = FALSE)

for (i in 1:(ncol(x)-1)) {
  for (j in 1:nrow(x)) {
    x[j,i] <-  substr(x[j,i], start=1, stop=7)
  }
}

for (i in 1:(ncol(x)-1)) {
  x[,i] <- as.numeric(x[,i])
}

for (i in ncol(x)) {
  for (j in 1:nrow(x)) {
    x[j,i] <-  substr(x[j,i], start=1, stop=7)
  }
}
print(
  xtable(
    align='lrrrrr', 
    digits=c(0, 2, 2, 1, 1, 3),
    x,
    caption="Response time model"), 
    caption.placement="top", 
  type='html')
```

# Error rates

```{r erdataplot1, results='asis', echo=F}
print(
  xtable(erdataplot1,
         caption='Condition means'),
  include.rownames=FALSE,
  caption.placement='top',
  type='html')
```

```{r tableErrorModelsimple, results='asis', cache=TRUE, echo=FALSE}
print(
  xtable(
    summary(
    ac.lmer.1
    )$coefficients,
    digits=c(0, 2, 2, 2, 5), 
    caption="Error rates model"),
  caption.placement="top", type='html')
```

<!-- ```{r} -->
<!-- rt.lmer.2 <- lmerTest::lmer( -->
<!--   data=rtdata, logrt ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ac.lmer.2 <- lme4::glmer( -->
<!--   data = erdata, error ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject),  -->
<!--   family="binomial", control = glmerControl(optimizer = "bobyqa")) -->
<!-- ``` -->

<!-- ```{r tableRTmodelMaximal, results='asis', echo=FALSE, cache=TRUE} -->
<!-- x = as.data.frame(format(summary(rt.lmer.2)$coef, scientific=FALSE), stringsAsFactors = FALSE) -->

<!-- for (i in 1:(ncol(x)-1)) { -->
<!--   for (j in 1:nrow(x)) { -->
<!--     x[j,i] <-  substr(x[j,i], start=1, stop=7) -->
<!--   } -->
<!-- } -->

<!-- for (i in 1:(ncol(x)-1)) { -->
<!--   x[,i] <- as.numeric(x[,i]) -->
<!-- } -->

<!-- for (i in ncol(x)) { -->
<!--   for (j in 1:nrow(x)) { -->
<!--     x[j,i] <-  substr(x[j,i], start=1, stop=7) -->
<!--   } -->
<!-- } -->
<!-- print( -->
<!--   xtable( -->
<!--     align='lrrrrr',  -->
<!--     digits=c(0,  2,2,2,2,3), -->
<!--     x, -->
<!--     caption="Maximal response time model"),  -->
<!--     caption.placement="top",  -->
<!--   type='html') -->
<!-- ``` -->

<!-- ```{r tableErrorModelMaximal, results='asis', cache=TRUE, echo=FALSE} -->
<!-- print( -->
<!--   xtable( -->
<!--     digits=3, summary( -->
<!--       ac.lmer.2)$coefficients, -->
<!--              caption="Maximal error rates model"), -->
<!--              caption.placement="top", type='html') -->
<!-- ``` -->