### 2016-09-25 experiments/e2/writeup/e2.Rmd
e2 in Rmd fomat for special issue - NEED TO DO STATS AND ADD TO TEX

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

```{r setup, include=FALSE, cache=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(pander)
```

```{r showInstructions, echo=FALSE, cache=TRUE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'word', 'number', 'word')
e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
i = data.frame(v, n, e)
names(i) = c("vagueness", "number", "example instruction")
pander(i, justify=c('center', 'left', 'left'))
```

```{r showItems, echo=FALSE, cache=TRUE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R') 
```

In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.

```{r getData, echo=T}
rawdata <- read.table('data.txt', sep=' ')
```

```{r}
tempdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
rtdata <- tempdata
bldata <- tempdata
```

This code creates the condition means with confidence intervals used for plotting:

```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log",        withinvars=c("Condition", "Vagueness", "Number", "Item"), idvar="Subject")
```

```{r rtplot, fig.width=5, fig.height=3}
dodge = position_dodge(width=0.2)
rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_log, x=Item, ymin=RT_log-ci, ymax=RT_log+ci,
                group=Condition, shape=Condition, fill=Condition)) +
  geom_line(position=dodge) +
  geom_errorbar(width=.2, position=dodge) +
  geom_point(size=3, position=dodge) +
  scale_shape_manual("",values = c(22, 22, 21, 21)) +
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), aspect.ratio=1,
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.background=element_rect(fill=NA, color='white')) +
  facet_grid(~Number) 
rtplot
```

```{r blplot, fig.width=5, fig.height=2.5}
bldata$Condition <- as.factor(paste(sep=':', bldata$Vagueness, bldata$Number))
bldata$response_category <- relevel(bldata$response_category, ref = 'expected')
blplot=ggplot(bldata) + 
  geom_bar(aes(Vagueness, fill=response_category), position = 'dodge') +
  #geom_errorbar() +
  facet_wrap(nrow=1,~Number) +
  scale_fill_grey(name="Response Type") +
  xlab(" ") +
  ggtitle("Response Type") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'))
blplot
```

<!-- ```{r} -->
<!-- dd$response_category <- relevel(dd$response_category, ref = 'expected') -->
<!-- barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5)) -->
<!-- ``` -->

<!-- , ymin=isBorderlineNormed-ci, ymax=isBorderlineNormed+ci,  group=Condition, fill=Condition)) + -->