### 2016-10-01 experiments/e2/writeup/e2-b-knitr.Rmd

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

```{setup, include=FALSE, cache=FALSE}
library(knitr)
```

```{loadFonts, message=FALSE, echo=FALSE}
library(fontcm)
library(extrafont)
#font_import()
#loadfonts()
```

```{loadLibs, message=FALSE, echo=FALSE}
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
```

```{preprocessing}
source('e2-preprocessing.R')
if( !file.exists('e2-b-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT)
  )
  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE)
}
dat <- read.delim('e2-b-data.txt')
dd <- read.delim('e2-b-data.txt')
```

```{showCompression, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Ratios for different numbers of dots in the arrays: smaller values are more discriminable. Blue is for the ratio between the smallest number in the array and the largest number in an array. Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array', echo=FALSE}
#  Show compression of ratios of numbers of dots in the items
discriminability = sort(unique(as.numeric(dd$discriminability)))
plot(y=c(1.5,1.5,1.5,1.5), 
     x=discriminability, 
     ylab='',
     xlab='',
     yaxt='n',
     type='n', 
     xaxt='n',
     axes=F)
lines(y=c(1.5,1.5,1.5,1.5), x=discriminability, type='o', pch=19)
text(y=c(1.15,1.15,1.15,1.15), x=discriminability, labels=round(discriminability,2),  cex=.7)
mtext('mean (ratio min/med,\nratio med/max)', side = 2, line = 1)
mtext('Discriminability index', side=1, line=3)
mtext('Discriminability for each item', side=3, line=1, cex=1.5)
axis(1)
box(which='plot', col='grey')
```

#Lmer model: before outlier removal

```{rmme}
load("data_processed.Rda")
```

```{lmer5, tidy=FALSE, cache=TRUE}
v5 <- lme4::lmer(data=dd,
                 RT_log ~
                   c_Vag + c_Num + c_Qty + c_Ord +
                   c_Num:c_Vag:c_Qty +
                   discriminability +
                   s_Trl +
                   RTprev_log +
                   nchar_instr +
                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{summaryV5, echo=FALSE}
summary(v5)
```

```{printXtableV5, echo=FALSE, results='asis'}
print(
  xtable(
    coef(summary(v5)
         ),
  caption = 'xtable v5'
  ),
  table.placement='htbp',
  latex.environments='center',
  size='small'
)
```

```{rSquaredInfo, echo=FALSE}
cat("R^2")
cor(fitted(v5), dd$RT_log)^2
```

```{plotModelCoefsAndCis, fig.width=5.5, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE}
# Plot model coefficients and ci's
dfr <- data.frame(coef = summary(v5)$coef[-1, 1], 
                     ci2.5  = confint(v5, method='Wald')[18:26, 1],
                     ci97.5 = confint(v5, method='Wald')[18:26, 2] )
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,10,1,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=range(dfr$ci2.5, dfr$ci97.5), 
     type='n', axes=F, xlab="", ylab="")
abline(v=0, lty=3, col='grey')
abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=3.5, cex.lab=1, las=1)
mtext("Model terms", side=2, line=8)
box(which='plot', col='grey')

```

```{plotLMERfnc, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='plotMLERfnc', echo=TRUE}
par(mfrow=c(2,4))
plotLMER.fnc(v5)

```

```{baayenPLots99, fig.width=6, fig.height=4, fig.cap='Baayen Model Criticism Plots', fig.pos='hbtp', echo=FALSE}
# Baayen 4-plot model criticism
par(mfrow=c(1,3), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))

```

#lmerTest Version}

```{lmerTestVersionOfModel, tidy=FALSE, cache=TRUE, echo=TRUE}
v6 <- lmerTest::lmer(data=dd,
                     RT_log ~
                       c_Vag + c_Num + c_Qty + c_Ord +
                       c_Num:c_Vag:c_Qty +
                       discriminability +
                       s_Trl +
                       RTprev_log +
                       nchar_instr +
                       (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{summaryLmerTest, cache=TRUE}
summary(v6)
```

#Lmer model: after outlier removal
not done yet.

#Borderline responses

```{barplotBorderline, fig.width=6, fig.height=4, echo=FALSE}
dd$response_category <- relevel(dd$response_category, ref = 'expected')
barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5))

```

```{tabBdl, echo=FALSE, results='asis'}
print(xtable(table(dd$response_category,dd$Vagueness),
             caption='Borderline cases counts'),
      table.placement='htbp',
      latex.environments='center',
      size='small')
```