### 2018-04-22 papers/2018-v-book/v-book.tex
2018: bare bones in Springer inhouse style

\documentclass[graybox,envcountchap,sectrefs%
,footinfo
%,referee
]{svmono}
% Native packages
\usepackage{mathptmx}
\usepackage{helvet}
\usepackage{courier}
\usepackage{type1cm}         
\usepackage{makeidx}		% allows index generation
\usepackage{graphicx}		% standard LaTeX graphics tool
\usepackage{multicol} 		% used for the two-column index
\usepackage[bottom]{footmisc}	% places footnotes at page bottom
% User packages
\usepackage{natbib}
\usepackage{lipsum}
\usepackage{placeins}
\makeindex             			% used for the subject index
                       				% please use the style svind.ist with
                       				% your makeindex program

\begin{document}

\author{Kees van Deemter and Matthew Green}
\title{Vagueness \& Rationality}
\subtitle{The Elusive Benefits of Vagueness: the Evidence So Far}
\maketitle

\frontmatter
\tableofcontents

% BibTeX users please use
\bibliographystyle{spbasic} %plain, harvard, apalike, chicago, astron, authordate, natbib.
\bibliography{bibliography}

\backmatter
\Extrachap{Glossary}
%\runinhead{glossary term} Write here the description of the glossary term. Write here the description of the glossary term. Write here the description of the glossary term.

\runinhead{natural language generation} the process of generating natural language from non-natural language input.

\printindex
\end{document}