### 2018-08-02 experiments/C_exp_1/archive/data_wrangling.Rmd

Get the data if file dat.Rda does not already exist.

```{r, get_the_data}
# if the output file dat.Rda already exists then load it, else do data wrangling
if( file.exists('dat.Rda') ) load('dat.Rda') else {

  # source required data wrangling functions
  source('functions/classifyResponse.R')
  source('functions/gatherData.R')

  # declare local variables
  number_of_valid_subjects <- 30
  number_of_rows <- 7680 # 7680
  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 256

    ## call user-defined functions

  # collect all individual subject files into one data frame
  dat <- gatherData(number_of_valid_subjects)

  # classify the response as expected, near, or far
  dat <- classifyResponse(dat)

   ## treat variables

  # SUBJECT
  # ensure Subject is a factor 
  dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))

  # TRIAL
  # Trial for subject, 1 to 256
  dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)

  # make a centred Trial for modeling
  dat$c_Trl <-dat$Trial - mean(dat$Trial)

  # make a scaled Trial for modelling
  dat$s_Trl <- as.numeric(scale(dat$Trial))

  # ID
  # id is a unique identifier for the 7680 row data
  dat$id <- factor(paste(paste(dat$Subject), 
                         paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

  # ITEM
  # create a centred numeric item variable for modeling
  dat$c_Itm <- ifelse(dat$Item==1, -.75, 
                      ifelse(dat$Item==2, -.25,
                             ifelse(dat$Item==3, .25, .75)))

  # make Item be a factor and assign labels
  dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))

  # add ratios for Item
  item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
          # 0.2500000 0.4705882 0.5909091 0.6666667
  item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
          # -1.3441995 -0.1316642  0.5297187  0.9461450
  item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
          # 0.5125000 0.6876471 0.7691558 0.8166667
  item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
          # -1.37582241 -0.06614191  0.54334858 0.89861574

  dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
  dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
  dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
  dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667

  dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
  dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
  dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
  dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450

  dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
  dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
  dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
  dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667

  dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
  dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
  dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
  dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574

  # VAGUENESS
  # Create a factor coding for Vagueness
  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
  dat$Vagueness <- as.factor(dat$Vagueness)

  # manually center Vagueness
  dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)

  # NUMBER
  # Create a factor coding for Number use
  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
  dat$Number <- as.factor(dat$Number)

  # manually center Number
  dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)

  # CONDITION
  # make a factor out of Condition, as f_Cnd
  dat$f_Cnd <- factor(dat$Condition, levels=c(1,2,3,4), labels=c('Vg:Nm', 'Cr:Nm', 'Vg:Vb', 'Cr:Vb'))

  # ORDER
  # give the levels of Order meaningful names
  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))

  # make a manually centred Order
  dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)

  # QUANTITY
  # give the levels of Quantity meaningful names
  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))

  # make a manually centred Quantity
  dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)

  # INSTRUCTION
  # add number of characters in the instruction # 29 30 34 36 38
  dat$nchar_instr = nchar(dat$Instruction)
  dat$nchar_instr_scaled = as.vector(scale(nchar(dat$Instruction), scale=TRUE))

  # make Instruction be a factor (17 levels)
  dat$Instruction <- as.factor(dat$Instruction)

  # RT
  # add transformations of RT
  dat$RT_RecSqd    <- 1/dat$RT^2
  dat$RT_Rec       <- 1/dat$RT
  dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
  dat$RT_log       <- log(dat$RT)
  dat$RT_sqt       <- sqrt(dat$RT)
  dat$RT_raw       <- dat$RT
  dat$RT_sqd       <- dat$RT^2

  # print to file a table with information about the design
  design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order, 
                                             Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
  design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
  row.names(design_info) <- NULL
  capture.output(  print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info.table")

  # put dat in better column order
  dat <- subset(dat, 
                select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, 
                           Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
                           c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
                           RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
                           Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
                           Instruction, nchar_instr
                ))

  # save dat as dat.Rda
  # This data set contains *all* trials 7680 including impossible trials and is mainly for graphs comparing different removals
  save(dat, file="dat.Rda")

  ####################################################
  # remove impossible trials from dat in new data dd #
  ####################################################

  # Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998 
  # i.e., throw out sticky fingers and timeouts, 
  # and the trials that followed sticky fingers and timeouts 
  # since they were likely affected by unusual previous trials.

  # lose impossible trials
  dd <- dat
  dd$RT[dd$RT == 1 ] <- NA
  dd$RT[dd$RT == 59998 ] <- NA
  dd <- dd[complete.cases(dd),]
  row.names(dd) <- NULL

  # add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
  dd$RTprev <- NA
  for (s in levels(dd$Subject)) {
    nrows = nrow(dd[dd$Subject==s,])
    for (i in 1:nrows) {
      if (i==1) {
        dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
      }
      else
        dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
    }
  }

  # add transformations of previous RT
  dd$RTprev_RecSqd    <- 1/dd$RTprev^2
  dd$RTprev_Rec       <- 1/dd$RTprev
  dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
  dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
  dd$RTprev_log       <- log(dd$RTprev)
  dd$RTprev_sqt       <- sqrt(dd$RTprev)
  dd$RTprev_raw       <- dd$RTprev
  dd$RTprev_sqd       <- dd$RTprev^2

  # number of higher-level cells in the design (abstracting over Order) = Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
  # number of measurements in eaach cell = 8
  # 332 * 8 = 256
  # 256 * 30 = 7680

  dd$measurement=0
  dd$cell=0
  for (s in levels(dd$Subject)) {
    cellcount=0
    for (i in levels(dd$Item)) {
      for (n in levels(dd$Number)) {
        for (v in levels(dd$Vagueness)) {
          for (q in levels(dd$Quantity)) {
            measurementcount=0
            cellcount=cellcount+1
            for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
              dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
              measurementcount=measurementcount+1
              dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
            }
          }
        }
      }
    }    
  }

  # put dd in better column order
  dd <- subset(dd, 
               select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, 
                          Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
                          c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
                          RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
                          RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
                          Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
                          Instruction, nchar_instr,
                          cell, measurement
               ))

  # save dd
  save(dd, file="dd.Rda")
}
```