### 2018-08-02 experiments/C_exp_1/archive/gelman_style_plot_regression_lines.Rmd

## Plot simple regressions on each individual predictor using log

```{r, fig.height=8, eval=TRUE, cache=FALSE}
par(mfrow=c(3,2))
subdata = subset(fulldata, subset=dfSubset== "RT>1&RT<59999")

# trial
fit.trial <- lm(subdata$log ~ subdata$Trial)
with(subdata, plot(log~Trial, cex=.5, col="grey", main="Trial"))
curve (coef(fit.trial)[1] + coef(fit.trial)[2]*x, add=TRUE)

# vagueness
fit.vagueness <- lm(subdata$log ~ subdata$c_Vag)
with(subdata, plot(log~jitter(c_Vag, 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Vagueness"))
axis(1, at=c(-.5,.5), labels=c("crisp","vague"))
curve (coef(fit.vagueness)[1] + coef(fit.vagueness)[2]*x, add=TRUE)

# number
fit.number <- lm(subdata$log ~ subdata$c_Num)
with(subdata, plot(log~jitter(c_Num, 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Number"))
axis(1, at=c(-.5,.5), labels=c("verbal","numeric"))
curve (coef(fit.number)[1] + coef(fit.number)[2]*x, add=TRUE)

# item
fit.item <- lm(subdata$log ~ subdata$c_Itm)
with(subdata, plot(log~jitter(c_Itm, 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Item"))
axis(1, at=c(-1.5, -.5, .5, 1.5), labels=levels(subdata$f_Itm))
curve (coef(fit.item)[1] + coef(fit.item)[2]*x, add=TRUE)

# order
fit.order <- lm(subdata$log ~ subdata$Order)
with(subdata, plot(log~jitter(as.numeric(Order), 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Order"))
axis(1, at=c(1,2), labels=c("small on left","small on right"))
curve (coef(fit.order)[1] + coef(fit.order)[2]*x, add=TRUE)

# quantity
fit.quantity <- lm(subdata$log ~ subdata$Quantity)
with(subdata, plot(log~jitter(as.numeric(Quantity), 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Quantity"))
axis(1, at=c(1,2), labels=c("choose small","choose large"))
curve (coef(fit.quantity)[1] + coef(fit.quantity)[2]*x, add=TRUE)
```