### 2018-08-02 experiments/C_exp_1/archive/plot_useful_plotting_code_for_using_later.Rmd

```{r}
with(rts, xyplot(data=rts, RecSqt ~ F1 | Effect, groups=F2, mins=mins, maxs=maxs,
                 par.settings = standard.theme(color = FALSE),
                 scales=list(x=list(relation="free",at=list( c(1,2),c(1,2),c(1,2),c(3,4),c(3,4),c(5,6) ))),
                 ylim=c(0.970, .984),
                 aspect = 1, layout=c(6,1), xlab="", 
                 panel=function(x,y,groups,subscripts,...){ 
                   panel.arrows(x,mins[subscripts],x,maxs[subscripts],...,lwd=1.5, angle=90, code=3, length=.01)
                   panel.xyplot(x,y,..., groups=groups, subscripts=subscripts, type='l', lty=3)
                   panel.xyplot(x,y,..., groups=groups, subscripts=subscripts, cex=1)
                   panel.abline(h=7.95, col="lightgrey")
                   draw.key(key=list(
                     text = list(text = levels(groups)[levels(groups)%in%groups[subscripts]]), 
                     points = list(pch =  trellis.par.get('superpose.symbol')$pch[1:nlevels(groups)][levels(groups)%in%groups[subscripts]])),
                     draw=TRUE , vp=grid::viewport(x=0.35,y=0.9))}) )
```