### 2018-08-18 experiments/D_exp_2/writeup/D_exp_2.Rmd
recompiled D-exp-2

---
title: "D_exp_2"
author: "Matt Green"
output:
  html_document:
    toc: true
    toc_float: false
---

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(broom.mixed)
library(lme4)
library(lmerTest)
library(dotwhisker)
library(kableExtra)
library(LMERConvenienceFunctions)
library(data.table)
library(gridExtra)
source("D_exp_2_functions/preprocessing.R")
source("D_exp_2_functions/pretty_coef_table.R")
source("D_exp_2_functions/summarySEwithin2.R")
root_dir <- getwd()
data_dir <- "D_exp_2_data"
```

```{r}
dat_full <- preprocessing(root_dir, data_dir)
dat <- subset(dat_full, subset=RT>0 & RT<25000) # loses 2 RTs that were judged to be outliers 25978 49871
dat_borderline <- dat
dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
dat$RT_log <- log(dat$RT)
```

# Info and Procedure (Method)

The main aim of this experiment was to see whether vague instructions would confer advantages over crisp alternatives when all instructions used numerals, and when there were vague and crisp versions of each of the comparison and matching strategies.

On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., "Choose a square with about 10 dots"). Participants then pressed a key to indicate that they had read the instruction.

The instruction remained on screen, and after 1000 ms, the arrays appeared (see Figure below).
<figure>
<img src="D_exp_2_screenshots/EXP_D_example_screenshot.bmp" style="width:480px;height:300px;">
</figure>

Response time was measured from the presentation of the arrays until the keypress indicating the participant's choice. The trial would timeout after 60 seconds if there was no response.

In this experiment, no feedback was given. This was because, in the vague conditions, we did not regard any response as "correct" or "incorrect", but instead as "expected response"; "borderline response"; and "extreme response", and we did not want to draw participants' attention to this distinction explicitly. Which choice the participant made was recorded for analysis.

# Full table of instructions

The instructions for this experiment were similar to C_exp_2, except that we changed "fewer than" to "far fewer than": this was because "far fewer than" can have borderline cases whereas "fewer than" cannot have borderline cases.

```{r, echo=FALSE}
instructions_table <- 
  dat %>%
    select(Item, Quantity, Vagueness, Selection, Instruction) %>%
    unique() %>%
    arrange(Item, Quantity, Vagueness, Selection) %>%
    spread(key=Vagueness, value=Instruction)
instructions_table %>% 
  kable(align='cccll', caption="Full table of instructions") %>% 
      kable_styling(full_width = F, position = "left", font_size = 11) 
```

# Means plots

```{r}
dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
```

```{r, "EXP_D_RT_means_plot", fig.width=7, fig.height=3, echo=FALSE}
dodge = position_dodge(width=0.2)
dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
ggplot(dat_plot, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, group=Condition, shape=Condition, fill=Condition)) +
  geom_line(position=dodge) +
  geom_errorbar(width=.2, position=dodge) +
  geom_point(size=2, position=dodge) +
  scale_shape_manual("",values = c(22, 22, 21, 21)) +
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), aspect.ratio=1,
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.background=element_rect(fill=NA, color='white')) +
  facet_grid(~Selection) 
```

```{r}
dat$Condition <- as.factor(paste(dat$Item, dat$Vagueness))
ggplot(data=dat, mapping=aes(x=Item, y=RT_log, group=Vagueness)) +
  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), aspect.ratio = 1) +
  facet_wrap(~Selection) +
  stat_summary(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), fun.y=mean, geom="line", lwd=2) +
  geom_smooth(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), method="loess")
```

```{r, "EXP_D_Borderline_response_distribution_plot", fig.width=7, fig.height=3.5, echo=FALSE}
dat_borderline$response_cat <- factor(dat_borderline$response_cat, levels= c("Near", "Expected", "Far"))
ggplot(dat_borderline) + 
  geom_bar(aes(response_cat, group=Selection:Vagueness, fill=Vagueness), position=position_dodge(width=NULL)) +   
  scale_fill_grey() + 
  xlab(NULL) + 
  ggtitle('Borderline response distribution') + 
  facet_grid(~Selection) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="top", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
```

# Hypotheses

We formulated the following hypotheses:

* [Hypothesis 1] Crisp/Vague RT:
    * Vague instructions should result in faster responses than crisp instructions.
* [Hypothesis 2] Comparison/Matching RT:
    * Instructions that allow comparison should result in faster responses than instructions that necessitate matching.
* [Hypothesis 3] Vagueness effect differs between comparison and matching: 
    * The effects of vagueness should differ when the selection task is comparison versus when it is matching.

# Results

Response times were trimmed at 25,000 ms, resulting in the loss of 2 trials that were judged to have outlier RTs. The remaining data were trimmed separately for each participant at 2.5 standard deviations below and above that participant's mean RT, leading to the loss of 204 trials (2.8% of the remaining trials). RTs in milliseconds were log-transformed so that they better approximated a normal distribution for the purposes of analysis.

## The first set of models

```{r, "rt_lmer_full_model_all_items", cache=TRUE}
dat_model <- dat
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
rtlmer = lmerTest::lmer(data=dat_model, RT_log ~ c_Vag * c_Sel + c_Itm + (1 + c_Vag * c_Sel + c_Itm | Subject))
pretty_coef_table(rtlmer, "rt_lmer_full_model_all_items")
```

```{r, "rt_lmer_for_comparison_all_items", cache=TRUE}
dat_model <- subset(dat, subset=Selection=='Comparison')
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
comp_lmer <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
pretty_coef_table(comp_lmer, "rt_lmer_for_comparison_all_items")
```

```{r, "rt_lmer_for_matching_all_items", cache=TRUE}
dat_model <- subset(dat, subset=Selection=='Matching')
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
match_lmer <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
pretty_coef_table(match_lmer, "rt_lmer_for_matching_all_items")
```

Considering the first set of models:

* [Hypothesis 1] Crisp/Vague RT:
    * _Vague instructions should result in faster responses than crisp instructions_ : Vague instructions resulted in faster responses than crisp instructions but the difference was not significant (p=0.716). 
* [Hypothesis 2] Comparison/Matching RT:
    * _Instructions that allow comparison should result in faster responses than instructions that necessitate matching_ : comparison instructions resulted in faster responses than matching instructions, and the difference was significant (p<0.001).
* [Hypothesis 3] Vagueness effect differs between comparison and matching
    * _The effects of vagueness should differ when the selection task is comparison versus when it is matching_ : Vagueness exerted effects in different directions for the two selection tasks (p<0.001). Separate analyses at each level of selection provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).

## The second set of models

The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018).

Testing for a main effect of X when X is a two-level factor and the model also includes a main effect of a three-level factor  and the interaction between X and Y.

Testing for a main effect of Vagueness when Vagueness is a two-level factor and the model also includes a main effect of a four-level factor Item and the interaction between Vagueness and Item.

```{r}
Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
dat$Item1 <- Item.numeric[1,]
dat$Item2 <- Item.numeric[2,]
dat$Item3 <- Item.numeric[3,]
m0 <- lm(RT_log ~ 
           Item1 + Vagueness:Item1 + 
           Item2 + Vagueness:Item2 + 
           Item3 + Vagueness:Item3 +
           Selection +
           Selection:Vagueness, 
         dat)
contrasts(dat$Item) <- contr.sum(4)
m.test <- lm(RT_log ~ 
               Item +
           Vagueness:Item1 + 
           Vagueness:Item2 + 
           Vagueness:Item3 +
           Selection +
           Selection:Vagueness, 
         dat)
m1 <- lm(RT_log ~ 
           Item1 + Vagueness:Item1 + 
           Item2 + Vagueness:Item2 + 
           Item3 + Vagueness:Item3 +
           Selection +
           Selection:Vagueness +
           Vagueness,
         dat)
m100 <- lm(RT_log ~
             Item +
             Vagueness +
             Item:Vagueness+
             Selection+
             Selection:Vagueness,
           dat)
#m1 and m100 are the same model
anova(m0,m1,test="F")
```

```{r}
Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
dat$Item1 <- Item.numeric[1,]
dat$Item2 <- Item.numeric[2,]
dat$Item3 <- Item.numeric[3,]
m3 <- lm(RT_log ~ 
           Item1 + 
           Item2 +  
           Item3 , 
         dat)
contrasts(dat$Item) <- contr.sum(4)
m4 <- lm(RT_log ~
           Item,
         dat)
# m3 and m4 are the same model
```

```{r}
Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
dat$Item1 <- Item.numeric[1,]
dat$Item2 <- Item.numeric[2,]
dat$Item3 <- Item.numeric[3,]

m.full <- lm(RT_log ~
             Item +
             Vagueness +
             Selection + 
             Item:Vagueness +
             Item:Selection + 
             Vagueness:Selection,
           dat)
m.reduced <- lm(RT_log ~
                Item1 + 
                Item2 + 
                Item3 + 
                Vagueness:Item1 + 
                Vagueness:Item2 +
                Vagueness:Item3 +
                Vagueness:Selection +
                Selection +
                Item:Selection
                ,dat  )
anova(m.reduced, m.full, test="F")
```

# Discussion