### 2018-08-31 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
consistent YAML

---
title: "A_pilot_1"
author: "Matt Green"
output:
  html_document:
    toc: yes
    theme: default
    css: styles.css
---

```{r load libraries, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(broom)
library(broom.mixed)
library(lme4)
library(lmerTest)
library(dotwhisker)
library(kableExtra)
library(LMERConvenienceFunctions)
library(data.table)
library(gridExtra)
source('summarySEwithin2.R')
source("A_pilot_1_functions/pretty_coef_table.R")
```

The experiment used subitizable numbers as well as bigger numbers. The instructions had the form: _Choose the square with {{few, many},{two, three, four, five, six, seven, eight, nine}} dots._ The variables were: subitizability; vagueness; target_size; gap_size; side (i.e., which side was the bigger number on?)

```{r "read the data from file, add some vars", message=FALSE}
# read in the data from file
din <- read_delim("data.txt", delim="\t")

# make an R object copy of that original data file
original_data <- din

# add observation number for the whole data set
din$obs <- 1:nrow(din)

# add trial index for each subject
din$trial <- 0
for (s in unique(din$subject)){
  din[din$subject==s, "trial"] <- 1:nrow(subset(din,subject==s))
}
```

```{r "make a tibble with factors explicit"}
# make a df to carry forward
d <- tibble(
  obs = din$obs, 
  trial = din$trial, 
  subject = as.factor(din$subject), 
  item = factor(din$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9")),
  gap = factor(din$gap, levels = c("BigGap", "SmallGap"), labels = c("Big Gap", "Small Gap")), 
  quan = factor(din$target, levels = c("BigTarget", "SmallTarget"), labels = c("Big Target", "Small Target")), 
  vag = factor(din$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
  subit = factor(ifelse(din$pair %in% c(1,2,5,6), "Subitizable", "Not subitizable")),
  side = factor(din$side), 
  text = factor(din$Text),
  err = 1 - din$isCorrect, 
  rt_ms = din$rt,
  rt_log = log( din$rt ),
  rt_reciprocal = -1/din$rt)
# reciprocal is rate of slowing: "We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' " - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662) 
```

```{r "make centred variables for factors"}
d$c.gap   <- ifelse(d$gap=="Big Gap", -.5, .5)
d$c.quan  <- ifelse(d$quan=="Big Target", -.5, .5)
d$c.vag   <- ifelse(d$vag=="Crisp", -.5, .5)
d$c.subit <- ifelse(d$subit=="Subitizable", -.5, .5)
d$c.side  <- ifelse(d$side=="Left", -.5, .5)
```

Size of the data is `r nrow(d)` trials.  There are `r nrow(subset(d,err==1))` error trials. Error trials are removed from analysis of RT

```{r, "remove error trials"}
d <- subset(d, err==0)
```

<!-- 1 was set to NA for being impossible ($-6~ms$). -->
<!-- The variable rt_ms has NA in the case that an error was made. The variable rt_raw includes rt for error trials. -->

## Eyeball the RTs

### Untrimmed

Notice min is -6; max in 20258.

```{r, "make untrimmed plot", fig.show='hide', warning=FALSE, echo=FALSE}
# all the (correct) RTs
summary(d$rt_ms)

data_for_plot <- d %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  arrange(rt_value)

ga1untrimmed = ggplot(data_for_plot, aes(x=rep(1:(nrow(data_for_plot)/3),times=3), y=rt_value)) + 
  geom_point(na.rm=TRUE) + 
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free_y") +
  xlab("observations, ordered by rt_value") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

ga2untrimmed = ggplot(data_for_plot, aes(rt_value)) +
  geom_density(na.rm=TRUE) +
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

myg1 <- grid.arrange(ga1untrimmed, ga2untrimmed, nrow=2, top = textGrob("Untrimmed", gp = gpar(fontsize=20, font=3)))
```

### Trimmed at 0 ms and 10,000 ms

```{r, "make trimmed plot", fig.show='hide', warning=FALSE, echo=FALSE}
# now with a simple cap imposed to select RTs between 0 and 10000 ms
summary(d$rt_ms[d$rt_ms >= 0 & d$rt_ms<10000])

data_for_plot <- d %>% 
  filter(rt_ms >= 0 & rt_ms <= 10000) %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  arrange(rt_value)

ga1trimmed = ggplot(data_for_plot, aes(x=rep(1:(nrow(data_for_plot)/3), times=3), y=rt_value)) + 
  geom_point(na.rm=TRUE) + 
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free_y") +
  xlab("observations, ordered by rt_value")+ theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

ga2trimmed = ggplot(data_for_plot, aes(rt_value)) +
  geom_density(na.rm=TRUE) +
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

myg2 <- grid.arrange(ga1trimmed, ga2trimmed, nrow=2, top = textGrob("Trimmed at 0 and 10,1000 ms", gp = gpar(fontsize=20, font=3)))
```

```{r, "put trimmed-untrimmed plot", fig.width=10,fig.height=10, echo=FALSE}
grid.arrange(myg1, myg2, nrow=2)
```

Select either trimmed or untrimmed data by comparing this plot with the untrimmed data. Also use this plot to choose the most normal data (logRT).

```{r selectTrim}
dat <- d %>% filter(rt_ms >= 0 & rt_ms <= 10000) 
```

```{r plot rt per trial by subject}
dat %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  filter(rt_metric=="rt_log") %>%
  ggplot(aes(x=trial, y=rt_value, col=vag)) +
  geom_point() +
  geom_smooth(se=F) +
  facet_wrap(~subject, ncol = 7) +
  theme(aspect.ratio = 1, panel.background = element_rect(fill="white", colour="black", linetype = "solid"))
```

```{r boxplot crisp and vague, fig.width=3, fig.height=3}
dat %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  filter(rt_metric=="rt_log") %>%
  ggplot(aes(group=vag, y=rt_value)) +
  geom_boxplot(aes(x=vag, col=vag)) + 
  theme(aspect.ratio=1)
```

# RT modelling

Do a first lmer model. You can't put subit or gap over items because they are between-items.

```{r, "lmer 1 compare different rt metrics using the same model structure", warning=FALSE, cache=TRUE}
lmer1.ms  <- lmer(data=dat, rt_ms ~         vag + subit + quan + gap + side + (1|subject) + (1|item))
lmer1.log <- lmer(data=dat, rt_log ~        vag + subit + quan + gap + side + (1|subject) + (1|item))
lmer1.rcp <- lmer(data=dat, rt_reciprocal ~ vag + subit + quan + gap + side + (1|subject) + (1|item))

models <- 
rbind(
  cbind(tidy(lmer1.ms),model="ms"), 
  cbind(tidy(lmer1.log),model="log"), 
  cbind(tidy(lmer1.rcp),model="rcp")
)

dwplot(models, vline = geom_vline(xintercept = 0)) + facet_wrap(~model, scales="free", ncol=1)
```

```{r lmer on logrt, warning=FALSE, cache=TRUE, fig.width=6, fig.height=2}
lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))

dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
  xlab("Coefficient") +
  ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)") + 
  theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8))

lmer2.rcp <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))

# lmer3 is closer to the original anova than lmer2
lmer3 <- lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item))

dwplot(lmer3 %>% tidy() %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
  xlab("Coefficient") + 
  ggtitle("rt_reciprocal ~ vag * subit + quan + gap + (vag+subit|subject) + (vag|item)") + 
  theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8))
```

<!-- ```{r} -->
<!-- lmer3b <- lme4::lmer(data=dat, logRT~subit*vag+quan+gap + (subit*vag|subject) + (vag|item))  -->
<!-- lmer3b_tidy <- tidy(lmer3b) -->
<!-- lmer3b_tidy$model="foo" -->
<!-- lmer3b_tidy %>% filter(group=="fixed") %>% dwplot() -->
<!-- ``` -->

<!-- ```{r, summaryLMER, results='asis'} -->
<!-- print(xtable(summary(lmer3)$coefficients, digits=3), type="html") -->
<!-- ``` -->

<!-- ## subitizable -->

<!-- ```{r, subitRT, results='asis'} -->
<!-- print(xtable(summary(lmer(data=dat, subset=subit=="Subitizable",logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html') -->
<!-- ``` -->

<!-- ## not subitizable -->

<!-- ```{r, notsubitrT, results='asis'} -->
<!-- print(xtable(summary(lmer(data=dat, subset=subit=="Not subitizable", logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html') -->
<!-- ``` -->

<!-- # Error rate analysis -->

<!-- ```{r errorrates, results='asis'} -->
<!-- print(xtable(summary( -->
<!-- 	glmer(data=d, err~vag*subit+quan+gap + (vag|subject) + (1|item),  -->
<!-- 				family=binomial,  -->
<!-- 				control = glmerControl(optimizer = "bobyqa") -->
<!-- 				) -->
<!-- 	)$coefficients, -->
<!-- 	digits=3),  -->
<!-- 	type='html') -->
<!-- ``` -->

<!-- # Plots of significant effects -->

<!-- ```{r plot_sig, fig.width=7.5, fig.height=3} -->
<!-- a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","subit","item"), idvar="subject") -->
<!-- as <- subset(a,subit=="Subitizable") -->
<!-- as$item <- as.character(as$item) -->
<!-- as$item[as$item=="7:3"] <- "3:7" -->
<!-- as$item <- factor(as$item) -->
<!-- pd <- position_dodge(.05) -->
<!-- plot1 = ggplot(data=as, aes(y=logRT, x=item, ymin=logRT-ci, ymax=ci+logRT, fill=vag, group=vag)) +  -->
<!--   geom_line(position=pd) + -->
<!--   geom_errorbar(width=.2, position=pd)+ -->
<!--   geom_point(pch=21, size=4, position=pd) + -->
<!--   scale_fill_grey(name=element_blank(),start=0,end=1)+ -->
<!--   theme_bw()+ -->
<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) + -->
<!--   ggtitle("Response time") -->

<!-- a <- summarySEwithin(data=d, measurevar="err", withinvars=c("vag","subit","item"), idvar="subject") -->
<!-- as <- subset(a,subit=="Subitizable") -->
<!-- as$item <- as.character(as$item) -->
<!-- as$item[as$item=="7:3"] <- "3:7" -->
<!-- as$item <- factor(as$item) -->
<!-- pd <- position_dodge(.05) -->
<!-- plot2 = ggplot(data=as, aes(y=err, x=item, ymin=err-ci, ymax=ci+err, group=vag, fill=vag))+ -->
<!--   geom_line(position=pd) + -->
<!--   geom_errorbar(width=.2, position=pd)+ -->
<!--   geom_point(pch=21, size=4, position=pd) + -->
<!--   scale_fill_grey(name=element_blank(),start=0,end=1)+ -->
<!--   theme_bw()+ -->
<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) + -->
<!--   ggtitle("Error rate") -->

<!-- grid.arrange(plot1, plot2, nrow=1, top="Subitizable only") -->
<!-- ``` -->

<!-- ```{r barcharts, fig.width=9, fig.height=3} -->
<!-- bwtheme <- standard.theme("pdf", color=FALSE) -->
<!-- b1=barchart(data=ddply(dat,.(vag,subit),summarise, logRT=mean(logRT)),  -->
<!-- 						logRT~subit,  -->
<!-- 						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme) -->
<!-- b2=barchart(data=ddply(d,.(vag,subit),summarise, Error_Rate=mean(err)),  -->
<!-- 						Error_Rate~subit,  -->
<!-- 						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme) -->
<!-- grid.arrange(b1,b2,ncol=2) -->
<!-- ``` -->

<!-- Subitizability x Vagueness x Quantity -->

<!-- ```{r plotMeansa, fig.width=7, fig.height=3} -->
<!-- a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","quan","subit"), idvar="subject") -->
<!-- pd <- position_dodge(.1) -->
<!-- 	ggplot( -->
<!-- 		a,  -->
<!-- 		aes(y=logRT, x=vag, group=quan, shape=quan, col=quan, ymin=logRT-ci, ymax=ci+logRT) -->
<!-- 	) +  -->
<!-- 	geom_point(size=4, position=pd) +  -->
<!-- 	geom_line(position=pd) +  -->
<!-- 	facet_grid(~subit) +  -->
<!-- 	geom_errorbar(width=.1, position=pd) +  -->
<!-- 	theme_bw() +  -->
<!-- 	theme(panel.grid.major=element_blank())  -->
<!-- ``` -->

<!-- ```{r qqplotLMER, fig.width=3, fig.height=3.5} -->
<!-- qqPlot(residuals(lmer2),cex=.5) -->
<!-- ``` -->