### 2018-08-31 experiments/B_pilot_2/preprocessing/preprocessing.Rmd
mvd writeup contents up

# Make data frame

```{r makeDataFrame}
dat <- read.delim("data_raw.txt", header=FALSE)
names(dat) <- c("condition", "leftN", "rightN", "vgn", "side", "string", "subject", "rt", "error")
dat$condition <- as.factor(dat$condition)
# item_ratio
for (c in unique(dat$condition)) {
  x=subset(dat, condition==c)
  x$item_ratio=unique(min(x$leftN,x$rightN)/max(x$leftN,x$rightN))
  dat[dat$condition==c, 'item_ratio']=x$item_ratio
}
# trial
for (s in unique(dat$subject)) {
  x=subset(dat, subject==s)
  x$trial=1:nrow(x)
  dat[dat$subject==s, 'trial']=x$trial
}
dat$subject <- as.factor(paste("sub",sep="",sprintf("%02d",dat$subject,sep="")))
dat$error <- as.logical(1-dat$error)
dat$Ord <- as.factor(ifelse(dat$leftN < dat$rightN, 'LtoR', 'RtoL'))
dat$c_Ord <- ifelse(dat$Ord=='LtoR', -.5, .5)
dat$bigside <- ifelse(dat$leftN > dat$rightN, 'L', 'R')
dat$Qty <- as.factor(ifelse(dat$side==dat$bigside, "Large", "Small"))
dat$bigside=NULL
dat$c_Qty <- ifelse(dat$Qty=='Small', -.5, .5)
dat$Gap <- as.factor(as.character(paste("gap", sep=" ", sprintf("%02d",abs(dat$leftN-dat$rightN)))))
dat$c_Gap[dat$Gap=="gap 05"] <- -1.5
dat$c_Gap[dat$Gap=="gap 10"] <- -.5
dat$c_Gap[dat$Gap=="gap 15"] <- .5
dat$c_Gap[dat$Gap=="gap 20"] <- 1.5
dat$Vgn <- as.factor(ifelse(dat$vgn=="P", "Crisp", "Vague"))
dat$vgn = NULL
dat$c_Vgn <- ifelse(dat$Vgn=='Crisp', -.5, .5)
dat$logrt <- log(dat$rt)
dat <- subset(dat, select=c(subject, trial, condition, leftN, rightN, item_ratio, side, string, Vgn, c_Vgn, Ord, c_Ord, Qty, c_Qty, Gap, c_Gap, error, rt, logrt))
row.names(dat) <- NULL
write.table(x=dat, "../data.txt")
```