### 2018-09-12 chapter-latex/our-approach-to-the-problem.tex

Are vague expressions processed more easily by readers than crisp ones? Like Lipman, we focus on situations where numerical information is used in order to identify a referent. Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \citep{vanDeemterCMR}. By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly. Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.

We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question. In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.

Let us start by explaining the task that was given to the participants in our experiments. We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities. In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared. Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality. The participant was asked to respond as quickly as possible while avoiding errors. We manipulated the instructions and the arrays in several ways across the four experiments.

All the experiments shared the following properties: Stimuli were created using the language GNU Octave \citep{eaton:2002} and the Psychophysics Toolbox extensions \citep{ptbx1, ptbx2}. The position of the dots was randomised per-trial. The order in which trials were presented was randomised per-participant. There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating. All participants self-reported fluency in English, and had normal, or corrected-to-normal vision. The experiment was conducted in a quiet room. Participants were asked to respond as quickly as possible while avoiding errors. There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room. All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \citep{lmerTest}. The complete data and analysis for this experiment and the others in this paper are available at \texttt{https://github.com/mjgreen/vagueness}.
%\texttt{\url{http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}

When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \citep{van123} and for processing Arabic numerals and number words \citep{Dehaene1996}. We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances. Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral. When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \citep{dehaene1993mental, gevers2006automatic}. We controlled which side the smaller number appeared on to avoid systematic influences of this effect.

There is abundant evidence \citep[e.g.,][]{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities. We performed a pilot experiment \citep{green2011costreduction} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper. We found that, when participants were confronted with a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}. The number of dots in an array is referred to as its cardinality. The physical arrangement of dots in each array is irregular.}, instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$. Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.

In a second pilot experiment \citep{green2013utility}, we again presented two dot arrays with the instruction in the form \emph{Choose the square with \ldots dots}. The arrays contained larger numbers of dots than in the first pilot: one array always contained 25 dots, and the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. Each stimulus can therefore be seen in terms of the numerical difference between the number of dots in one array and the number of dots in the other: giving numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays. Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}. Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}. We found, as expected, that responses were faster and more accurate for vague instructions than for crisp instructions. We also found an interaction between vagueness and numerical distance such that there were diminishing returns for vagueness as numerical distance increased, until, at the biggest numerical distance, there was no real difference any more between crisp and vague instructions. We interpreted this pattern as showing that cognitive load is relatively easy in both conditions when the stimuli are most discriminable, and so vagueness confers no additional advantage in those most easily discriminable stimuli.

However, the picture painted by these findings from the second pilot experiment might be misleading. First of all, there is a possibly confounding factor. Contrast an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'. One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. Since these two differences could not be separated in that experiment, the finding of a vagueness advantage is vulnerable to an alternative interpretation, that the crisp-vague difference was really an advantage for the verbal form of the quantifier. In our next experiment (reported below as Experiment 1) we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare the crisp-vague difference and the numeric-verbal difference in the same experiment.

In Experiment 1, reported below, we also addressed another issue with the second pilot experiment: participants chose one of two dot arrays - therefore the `vague' quantifiers (few and many) uniquely identified one square. Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise''. The quantifiers few and many might not have realised their potential for vagueness in the context of a choice between (only) two alternatives - where there is no borderline alternative that could result in few and many being "not well-defined". Furthermore using definite articles in the instructions may have reinforced the participant's impression that only one choice counted as correct, and this impression could have been reinforced by the use of error feedback.

In the experiments we report below (Experiments 1, 2 and 3) we used what we learned from the pilot experiments to design situations that we believe did address the difference between crisp and vague instructions while taking account of the most important alternative explanations of this crisp-vague difference.