### 2018-10-27 manuscript/v-book-2018.tex

\documentclass[ %
  graybox       %
 ,envcountchap  %
 ,sectrefs      %
% ,footinfo      %
% ,graphics      %
%,referee
]{svmono}
% Native packages
\usepackage{mathptmx}
\usepackage{helvet}
\usepackage{courier}
\usepackage{type1cm}
\usepackage{makeidx}		    % allows index generation
\usepackage{graphicx}		    % standard LaTeX graphics tool
\usepackage{multicol} 		    % used for the two-column index
\usepackage[bottom]{footmisc}	% places footnotes at page bottom
% User packages
\usepackage{natbib}
\usepackage{lipsum}
\usepackage{placeins}
\usepackage{multirow}			% added by Matt
%\makeindex             		% used for the subject index: please use the style svind.ist with your makeindex program
\begin{document}
\author{Matthew James Green and Kees van Deemter}
\title{Vagueness \& Rationality}
\subtitle{The Elusive Benefits of Vagueness: the Evidence So Far}
\date{}
\maketitle
\frontmatter
\tableofcontents
\mainmatter
\chapter{The Elusive Benefits of Vagueness: Evidence from Experiments}\label{chapterlabel}
\input{abstract}
\section{Introduction}\label{introduction}\input{introduction}
\section{Related work}\label{related-work}\input{related-work}
\section{Our Approach to the Problem}\label{our-approach-to-the-problem}\input{our-approach-to-the-problem}
\section{Experiment 1: separating vagueness from instruction format}\label{e1}\input{e1}
\section{Experiment 2: focus on instructions that contain numerals}\label{e2}\input{e2}
\section{Experiment 3: focus on instructions that do not contain numerals}\label{e3}\input{e3}
\section{Discussion of experiments 2 and 3}\label{discussion-of-e2-and-e3}\input{discussion-of-e2-and-e3}
\section{General Discussion}\label{general-discussion}\input{general-discussion}
\backmatter
\bibliographystyle{spbasic}
\bibliography{v-book-2018}
\backmatter
%\printindex
\end{document}